# Analysis {#analysis}

While **this chatper has not been written**, a few resources and basic examples were made available to help out until this chapter is written.

## Overview

## dplyr {#dplyr}

Using `sparklyr`, you can apply the same data analysis techniques described in [Chapter 5 - Data transformation - R for Data Science](http://r4ds.had.co.nz/transform.html) by Garrett Grolemund and Hadley Wickham. 

Once you understand `dplyr`, you can make use of `dplyr` and `sparklyr` as follows:

```{r warning=FALSE, message=FALSE}
library(sparklyr)
library(dplyr)

# Connect to Spark
sc <- spark_connect(master = "local")

# Use dplyr's copy_to() to copy the iris dataset to Spark
iris_tbl <- copy_to(sc, iris, overwrite = TRUE)

# The iris_tbl is a Spark data frame compatible with dplyr
iris_tbl
```

```{r warning=FALSE}
# Transform iris_tbl with dplyr as usual
iris_tbl %>% 
  group_by(Species) %>% 
  summarise_all(funs(mean))
```

### Pass-through Functions

```{r eval=FALSE}
dates_tbl <- copy_to(sc, data.frame(a = "1/2/2001"))
dates_tbl %>% select(b = to_date(a)) %>% as.sql()
```

See [Spark SQL built-in functions](https://spark.apache.org/docs/2.3.0/api/sql/index.html).

### Resources {#dplyr-reosurces}

To understand `dplyr` further, I would recommend taking a look at the following vignettes:

- [Introduction to dplyr](https://cran.r-project.org/web/packages/dplyr/vignettes/dplyr.html)
- [Two-table verbs](https://cran.r-project.org/web/packages/dplyr/vignettes/two-table.html)
- [Window functions](https://cran.r-project.org/web/packages/dplyr/vignettes/window-functions.html)
- [Programming with dplyr](https://cran.r-project.org/web/packages/dplyr/vignettes/programming.html)

## DBI {#dbi}

The `DBI` provides an database interface for R, meaning, if you are familiar with SQL, you can make use of `DBI` to perform SQL queries in Spark using `sparklyr`. To learn more about `DBI`, I would recommend reading first [A Common Database Interface (DBI)](https://cran.r-project.org/web/packages/DBI/vignettes/DBI-1.html). Once you are familiar with `DBI`, you can use this package with `sparklyr` as follows:

```{r warning=FALSE, message=FALSE}
library(DBI)

dbGetQuery(sc,
  "SELECT mean(Sepal_Length), mean(Sepal_Width), 
          mean(Petal_Length), mean(Petal_Width)
   FROM iris
   GROUP BY Species")
```

More advanced `DBI` resources are available in the following vignettes:

- [A Common Interface to Relational Databases from R and S – A Proposal](https://cran.r-project.org/web/packages/DBI/vignettes/DBI-proposal.html)
- [Implementing a new backend](https://cran.r-project.org/web/packages/DBI/vignettes/backend.html)
- [DBI specification](https://cran.r-project.org/web/packages/DBI/vignettes/spec.html)
