# Data {#data}

While **this chatper has not been written.**, a few resources are available to help explore these topics until this chapter gets written.

## Overview

## Formats

```
spark_read_csv()
spark_read_jdbc()
spark_read_json()
spark_read_libsvm()
spark_read_orc()
spark_read_parquet()
spark_read_source()
spark_read_table()
spark_read_text()
```

## External

### Cassandra

See [https://blog.rstudio.com/2017/07/31/sparklyr-0-6/#external-data-sources](https://blog.rstudio.com/2017/07/31/sparklyr-0-6/#external-data-sources).

### Databases

See [https://blog.rstudio.com/2017/07/31/sparklyr-0-6/#external-data-sources](https://blog.rstudio.com/2017/07/31/sparklyr-0-6/#external-data-sources).

### HBase

## Troubleshooting

### Troubleshoot CSVs

```{r eval=FALSE}
writeLines(c("bad", 1, 2, 3, "broken"), "tmp/bad.csv")
```

There are a couple modes that can help troubleshoot parsing issues:
- **PERMISSIVE**: `NULL`s are inserted for missing tokens.
- **DROPMALFORMED**: Drops lines which are malformed.
- **FAILFAST**: Aborts if encounters any malformed line.

Which can be used as follows:

```{r=FALSE}
spark_read_csv(
  sc,
  "bad",
  "tmp/bad.csv",
  columns = list(foo = "integer"),
  infer_schema = FALSE,
  options = list(mode = "DROPMALFORMED"))
```
```
# Source:   table<bad> [?? x 1]
# Database: spark_connection
    foo
  <int>
1     1
2     2
3     3
```

In Spark 2.X, there is also a secret column `_corrupt_record` that can be used to output those incorrect records:

```{r eval=FALSE}
spark_read_csv(
  sc,
  "decimals",
  "tmp/bad.csv",
  columns = list(foo = "integer", "_corrupt_record" = "character"),
  infer_schema = FALSE,
  options = list(mode = "PERMISIVE")
)
```
```
# Source:   table<decimals> [?? x 2]
# Database: spark_connection
    foo `_corrupt_record`   
  <int> <chr>               
1     1 NA                  
2     2 NA                  
3     3 NA                  
4    NA sdfsdfds            
5    NA 2.16027303300001e+31
```

### Column Names

By default, `sparklyr` sanitizes column names by translating characters like `.` to `_`, this was required in Spark 1.6.X to avoid couple nuances in Spark. However, to disable this functionality, you can run the following code:

```{r eval=FALSE}
options(sparklyr.sanitize.column.names = FALSE)
dplyr::copy_to(sc, iris, overwrite = TRUE)
```
```
# Source:   table<iris> [?? x 5]
# Database: spark_connection
   Sepal.Length Sepal.Width Petal.Length Petal.Width Species
          <dbl>       <dbl>        <dbl>       <dbl> <chr>  
 1          5.1         3.5          1.4         0.2 setosa 
 2          4.9         3            1.4         0.2 setosa 
 3          4.7         3.2          1.3         0.2 setosa 
 4          4.6         3.1          1.5         0.2 setosa 
 5          5           3.6          1.4         0.2 setosa 
 6          5.4         3.9          1.7         0.4 setosa 
 7          4.6         3.4          1.4         0.3 setosa 
 8          5           3.4          1.5         0.2 setosa 
 9          4.4         2.9          1.4         0.2 setosa 
10          4.9         3.1          1.5         0.1 setosa 
# ... with more rows
```

