```{r analysis-setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# Analysis {#analysis}

## Introduction

Previous chapters focused on introducing Spark, R and helping you get started with the tools you need throughout this book. In this chapter you will learn how to do data analysis in Spark from R. Data analysis may become the most common task you will do when working with Spark. This chapter will serve as a foundation to later chapters because concepts from this chapter will apply to properly prepare data when doing modeling, graph processing, streaming and other related topics that might not be strictly considered data analysis.

For those who are new to R, this might be a good time to consider complementing this chapter with R for Data Science [@intro-r-for-data-science] and other online resources to learn R. This chapter will try to briefly introduce all the concepts it presents.

### Searching for insights

In a data analysis project, the main goal is to search for insights that can be derived from the data.  It is the results of the data transformations identified during the data analysis phase that can later be formalized into artifacts such as dashboard or model pipelines. The output of the data transformations could be models, aggregations or visualizations. Most data analysis projects follow a set of steps outlined in *Figure 3.1*. 

```{r echo=FALSE, out.width='100%', out.height='220pt', fig.cap='General steps of a data analysis', fig.align = 'center'}
nomnoml::nomnoml("
#direction: right
#edgeMargin: 4
#padding: 25
#fontSize: 18

[Import] -> [Understand]
[Understand |
  [Wrangle] -> [Visualize] 
  [Visualize] -> [Model]
  [Model] -> [Wrangle]
]
[Understand] -> [Communicate]
")
```

### R as an interface to Spark

For data analysis, the ideal approach  is to let Spark do what its good at. It excels at being a parallel computation engine that works at a large scale.  Spark goes beyond offering generic calculations. Out of the box, Spark includes libraries that actually can do a lot of what analysts usually do in R, but for large amounts of data. *Figure 3.2* paraphrases the four main capabilities available to data analysts in Spark.

```{r echo=FALSE, out.width='100%', out.height='220pt', fig.cap='Spark capabilities', fig.align = 'center', eval = TRUE}
nomnoml::nomnoml("
#direction: down
#fontSize: 18
#padding: 30
[Apache Spark |
  [SQL Engine]
  [Machine Learning]
  [Graph analysis]
  [Streaming]
]
")
```

Thanks to Spark's libraries, most of the Data Science project steps can be completed inside Spark.  For example, selecting, transforming and modeling can all be done by Spark.  The idea is to use R to tell Spark what data operations to run (import, wrangle, model), and then focus on only bringing back into R the results of the operation.  

```{r echo=FALSE, out.width='100%', out.height='220pt', fig.cap='R as an interface for Spark', fig.align = 'center', eval = TRUE}
nomnoml::nomnoml("
#direction: right
#padding: 40
#spacing: 140
#arrowSize:0.2
#.tag: visual=none
[<tag>R]Collect results->Push compute[Spark]
[Spark]->[R]
")
```

The `sparklyr` package focuses on implementing the principle mentioned in the previous section.  Most of its functions are mainly wrappers on top of Spark API calls.

The idea is take advantage of Spark's analysis components instead of R's.  For example, if the analyst needs to fit a Linear Regression model, instead of using the familiar `lm()` function, for data available via Spark, the analyst would use the `ml_linear_regression()` function.  The R function will actually run Scala code that runs the Spark's API model.  

```{r echo=FALSE, out.width='100%', out.height='220pt', fig.cap='R as an interface for Spark', fig.align = 'center', eval = TRUE}
nomnoml::nomnoml("
#direction: right
#padding: 20
#spacing: 140
#arrowSize:0.2
[R/sparklyr | ml_linear_regression()]Wraps code ->  [Scala |
        	var lr = new LinearRegression()
            var lrModel = lr.fit(training)]
 
")
```

For more common data manipulation tasks, `sparklyr` provides a back-end for `dplyr`.  This means that already familiar `dplyr` verbs can be used in R, and then `sparklyr` and `dplyr` will translate those actions into Spark SQL statements, see *figure 3.8*.

```{r echo=FALSE, out.width='100%', out.height='220pt', fig.cap='R as an interface for Spark', fig.align = 'center', eval = TRUE}
nomnoml::nomnoml("
#direction: right
#padding: 20
#spacing: 140
#arrowSize:0.2
[R/dplyr | filter(y == 1)]Wraps code ->  [SQL |
        	... WHERE y = 1 ...]
 
")
```

## Exercise

In order to practice as you learn, the rest of this chapter's code will use a single exercise that runs in the **local** Spark master, this allows the code to be replicated in your laptop. Please, make sure to already have `sparklyr` and a local copy of Spark installed. The installation of Spark can be done by using the utility that comes with the package.  For more information on how to do that please see the Local section in the [Connections](conections/#connections-local) chapter.

First, load the `sparklyr` and `dplyr` packages, and open a new **local** connection.
```{r}
library(sparklyr)
sc <- spark_connect(master = "local")
```



## Import / Access

The step of importing data is to be approached differently when using Spark with R, as oppose to R alone.  When used alone, importing data means that R will read files and import the information into memory  But when used with Spark, it is important to only focus in importing results into R.  The data is either imported or accessed by Spark. This way, the actual analysis takes place inside the Spark session. 
```{r echo=FALSE, out.width='100%', out.height='220pt', fig.cap='Import Data to Spark not R', fig.align = 'center', eval = TRUE}
nomnoml::nomnoml("
#direction: right
#padding: 40
#spacing: 140
#arrowSize:0.2
#.tag: visual=none
[<tag>R]Collect results->Push compute[Spark]
[Spark]->[R]
[Spark]<-Import/Access[Data;Source]
")
```

An example of accessing versus importing is found in an enterprise environment.  More likely, Spark sessions are created on top of Hadoop clusters, so data would already be available to be accessed directly by Spark via either a Hive table, or through the Hadoop File System (HDFS). 

The decision of having Spark either access the data source or to import data into memory is mostly a decisions based on speed and performance.  That will be covered in the [Tuning](tuning) chapter.

The exercise's Spark session does not have any data.  So the next step is to prime the session with data, in this case `mtcars`. The `copy_to()` command from `dplyr` can be used for that. The mechanics of this operation is explained in the [Getting Started](starting.html#starting-sparklyr-hello-world) chapter. 

```{r}
library(dplyr)
cars <- copy_to(sc, mtcars, "mtcars_remote")
```

In an enterprise setting, `copy_to()` should only be used to transfer small tables from R, such as a look up value table.

## Wrangle

### Transformations using `dplyr`

The goal is to keep is much of the data transformations in R code as possible. To accomplish that, instead of writing SQL statements to perform data exploration, it is better to take advantage of `sparklyr`'s `dplyr` back end interface.  

In the R environment, *cars* can be treated as if it is a local data frame, so `dplyr` verbs can be used, and in a piped fashion.

```{r}
cars %>%
  group_by(am) %>%
  summarise(mpg_mean = mean(mpg, na.rm = TRUE))
```
```
## # Source: spark<?> [?? x 2]
##      am mpg_mean
##   <dbl>    <dbl>
## 1     0     17.1
## 2     1     24.4
```

Instead of importing the *mtcars_remote* data set from Spark, and then performing the aggregation, `dplyr` converts the verbs into SQL statements that are then sent to Spark.  The `show_query()` command makes it possible to peer into the SQL statement that `sparklyr` and `dplyr` created and sent to Spark.

```{r}
cars %>%
  group_by(am) %>%
  summarise(mpg_mean = mean(mpg, na.rm = TRUE)) %>%
  show_query()
```
```
## <SQL>
## SELECT `am`, AVG(`mpg`) AS `mpg_mean`
## FROM `mtcars_remote`
## GROUP BY `am`
```
As it is evident, it will not be necessary to have to see the resulting query every time `dplyr` verbs are being used. The focus can remain on obtaining insights from the data, as opposed to figuring out how to express a given set of transformation in SQL.

```{r}
cars %>%
  group_by(am) %>%
  summarise(
    wt_mean = mean(wt, na.rm = TRUE),
    mpg_mean = mean(mpg, na.rm = TRUE)
  )
```
```
## # Source: spark<?> [?? x 3]
##      am wt_mean mpg_mean
##   <dbl>   <dbl>    <dbl>
## 1     0    3.77     17.1
## 2     1    2.41     24.4
```

Most of the data transformation made available by `dplyr` to work with local data frames are also available to use with a Spark connection.  This means that a general approach to learning `dplyr` can be taken in order to gain more proficiency with data exploration and preparation with Spark.  The chapter on Data Transformation in the R for Data Science [@intro-r-for-data-science] book should be a great help with this.  If proficiency with `dplyr` is not an issue for the reader, please take some time to experiment with different `dplyr` functions against the *cars* table.

### Correlations

A very common exploration technique is to calculate and visualize correlations.  The Spark API provides an internal function that calculates correlations across the entire data set.  The results are then returned to R as a `data.frame` object. 

```{r}
ml_corr(cars) %>%
  as_tibble()
```

```
## # A tibble: 11 x 11
##       mpg    cyl   disp     hp    drat     wt    qsec
##     <dbl>  <dbl>  <dbl>  <dbl>   <dbl>  <dbl>   <dbl>
##  1  1     -0.852 -0.848 -0.776  0.681  -0.868  0.419 
##  2 -0.852  1      0.902  0.832 -0.700   0.782 -0.591 
##  3 -0.848  0.902  1      0.791 -0.710   0.888 -0.434 
##  4 -0.776  0.832  0.791  1     -0.449   0.659 -0.708 
##  5  0.681 -0.700 -0.710 -0.449  1      -0.712  0.0912
##  6 -0.868  0.782  0.888  0.659 -0.712   1     -0.175 
##  7  0.419 -0.591 -0.434 -0.708  0.0912 -0.175  1     
##  8  0.664 -0.811 -0.710 -0.723  0.440  -0.555  0.745 
##  9  0.600 -0.523 -0.591 -0.243  0.713  -0.692 -0.230 
## 10  0.480 -0.493 -0.556 -0.126  0.700  -0.583 -0.213 
## 11 -0.551  0.527  0.395  0.750 -0.0908  0.428 -0.656 
## # ... with 4 more variables: vs <dbl>, am <dbl>,
## #   gear <dbl>, carb <dbl>
```

The `corrr` R package specializes in correlations.  It contains friendly functions to prepare and visualize the results.  Included inside the package is a back-end for `sparklyr` table objects, so it will not return an error if a non local table is passed to it.  In the background, the `correlate()` function runs `ml_corr()`, so there is no need to collect any data into R prior running the command.

```{r}
library(corrr)

cars %>%
  correlate(use = "pairwise.complete.obs", method = "pearson") 
```

```
## # A tibble: 11 x 12
##    rowname     mpg     cyl    disp      hp     drat      wt
##    <chr>     <dbl>   <dbl>   <dbl>   <dbl>    <dbl>   <dbl>
##  1 mpg      NA      -0.852  -0.848  -0.776   0.681   -0.868
##  2 cyl      -0.852  NA       0.902   0.832  -0.700    0.782
##  3 disp     -0.848   0.902  NA       0.791  -0.710    0.888
##  4 hp       -0.776   0.832   0.791  NA      -0.449    0.659
##  5 drat      0.681  -0.700  -0.710  -0.449  NA       -0.712
##  6 wt       -0.868   0.782   0.888   0.659  -0.712   NA    
##  7 qsec      0.419  -0.591  -0.434  -0.708   0.0912  -0.175
##  8 vs        0.664  -0.811  -0.710  -0.723   0.440   -0.555
##  9 am        0.600  -0.523  -0.591  -0.243   0.713   -0.692
## 10 gear      0.480  -0.493  -0.556  -0.126   0.700   -0.583
## 11 carb     -0.551   0.527   0.395   0.750  -0.0908   0.428
## # ... with 5 more variables: qsec <dbl>, vs <dbl>,
## #   am <dbl>, gear <dbl>, carb <dbl>
```

The `correlate()` function returns a local R object that `corrr` recognizes. This way, it is easy to perform more functions on top of the results. In this case, the `shave()` command turns all of the duplicated results into `NA`'s

```{r}
cars %>%
  correlate(use = "pairwise.complete.obs", method = "pearson") %>%
  shave()
```

```
## # A tibble: 11 x 12
##    rowname     mpg     cyl    disp      hp     drat      wt
##    <chr>     <dbl>   <dbl>   <dbl>   <dbl>    <dbl>   <dbl>
##  1 mpg      NA      NA      NA      NA      NA       NA    
##  2 cyl      -0.852  NA      NA      NA      NA       NA    
##  3 disp     -0.848   0.902  NA      NA      NA       NA    
##  4 hp       -0.776   0.832   0.791  NA      NA       NA    
##  5 drat      0.681  -0.700  -0.710  -0.449  NA       NA    
##  6 wt       -0.868   0.782   0.888   0.659  -0.712   NA    
##  7 qsec      0.419  -0.591  -0.434  -0.708   0.0912  -0.175
##  8 vs        0.664  -0.811  -0.710  -0.723   0.440   -0.555
##  9 am        0.600  -0.523  -0.591  -0.243   0.713   -0.692
## 10 gear      0.480  -0.493  -0.556  -0.126   0.700   -0.583
## 11 carb     -0.551   0.527   0.395   0.750  -0.0908   0.428
## # ... with 5 more variables: qsec <dbl>, vs <dbl>,
## #   am <dbl>, gear <dbl>, carb <dbl>
```
Finally, the results can be easily visualized using `rplot()`.  This function returns a `ggplot` object.

```{r}
cars %>%
  correlate(use = "pairwise.complete.obs", method = "pearson") %>%
  shave() %>%
  rplot()
```

```{r, eval = TRUE, out.width='800pt', out.height='400pt', fig.align = 'center', echo = FALSE}
render_image("images/analysis-wrangling-1.png")
```

## Visualize

Visualizations are fundamentally a human task.  They are a vital tool to help us find patterns from data.  It is easier for us to identify outliers in a data set of 1,000 observations when plotted then when reading them from a list.

R is great at data visualizations. Its capabilities for creating plots is extended by the many R packages that focus on this analysis step.   Unfortunately, the vast majority of R functions that create plots depend on the data already being in local memory within R, so they fail when using a remote table inside Spark. 

It is possible to create visualizations in R from data source from Spark. To understand how to do this, let's first break down how computer programs build plots:

```{r echo=FALSE, out.width='100%', out.height='100pt', fig.cap='Stages of a plot', fig.align = 'center', eval = TRUE}
nomnoml::nomnoml("
#direction: right
#edgeMargin: 4
#padding: 25
#fontSize: 18

[Raw data] -> [Transformed Data]
[Transformed Data] -> [Map data to coordinates]
[Map data to coordinates] -> [Draw plot]
")
```

For example, to create a bar plot in R, we simply call a function:

```{r, eval = TRUE, out.width='500pt', out.height='400pt', fig.cap='Plotting inside R', fig.align = 'center', eval = TRUE}
ggplot(aes(as.factor(cyl), mpg), data = mtcars) + geom_col()
```

In this case, the `mtcars` raw data was **automatically** transformed into three discrete aggregated numbers, then each result was mapped into an `x` and `y` plane, and then the plot was drawn, see *figure 3.8*. As R users, all of the stages of building the plot are conveniently abstracted for us.

```{r echo=FALSE, out.width='100%', out.height='200pt', fig.cap='R plotting function', fig.align = 'center', eval = TRUE}
nomnoml::nomnoml("
#direction: right
#edgeMargin: 4
#padding: 25
#fontSize: 18

[R |
  [mtcars] -> [Total mpg for each cyl]
  [Total mpg for each cyl] -> [x = cyl, y = mpg]
  [x = cyl, y = mpg] -> [Draw plot]
]
")
```

### Recommended approach

In essence, the approach for visualizing is the same as in wrangling.  Push the computation to Spark, and then collect the results in R for plotting.  The heavy lifting of preparing the data, such as in aggregating the data by groups or bins, can be done inside Spark, and then collect the much smaller data set into R.  Inside R, the plot becomes a more basic operation.  For example, to plot a histogram, the bins are calculated in Spark, and then in R, the plot that will be used is a simple column plot, as opposed to a histogram plot, because there is no need for R to re-calculate the bins.


```{r echo=FALSE, out.width='100%', out.height='200pt', fig.cap='Plotting with Spark & R', fig.align = 'center', eval = TRUE}
nomnoml::nomnoml("
#direction: right
#edgeMargin: 4
#padding: 25
#fontSize: 18
[Spark |
  [Raw data] -> [Transformed Data]
]
[R |
  [Map to coordinates] -> [Draw plot]
]
[Spark] -> [R]
")
```

### Simple plots

There are a couple of key steps when codifying the "Transform remotely, plot locally" approach.  First, ensure that the transformation operations happen inside Spark. In the example below, `group_by()` and `summarise()` will run as SQL inside the Spark session.  The second is to bring the results back into R after the data has been transformed.  Make sure to transform and then collect, in that order, because if `collect()` is run first, then all R will try to ingest the entire data set from Spark. Depending on the size of the data, collecting all of the data will slow down or may even bring down your system.

```{r}
car_group <- cars %>%
  group_by(cyl) %>%
  summarise(mpg = sum(mpg, na.rm = TRUE)) %>%
  collect()

car_group
```

```
## # A tibble: 3 x 2
##     cyl   mpg
##   <dbl> <dbl>
## 1     6  138.
## 2     4  293.
## 3     8  211.
```

In the example, now that the data has been pre-aggregated and collected into R, only three records are passed to the plotting function.

```{r}
ggplot(aes(as.factor(cyl), mpg), data = car_group) + geom_col()
```

```{r, eval = TRUE, fig.width=10, fig.height=5, fig.cap='Plot from Spark', fig.align = 'center', echo = FALSE}
render_image("images/analysis-visualizations-1.png")
```

Thanks to the consistency among the `tidyverse` packages, the entire operation can be written in a single piped code segment:

```{r}
cars %>%
  group_by(cyl) %>%
  summarise(mpg = sum(mpg, na.rm = TRUE)) %>%
  collect() %>%
  ggplot(aes(as.factor(cyl), mpg)) + 
  geom_col()
```

```{r, eval = TRUE, fig.width=10, fig.height=5, fig.cap='Plot from Spark', fig.align = 'center', echo = FALSE}
render_image("images/analysis-visualizations-1.png")
```

Using this approach, most visualizations can be easily produced. 

### Histograms

There are plots that are both useful and commonly used, but their calculations are not easily reproducible. For example, producing a histogram that runs over the entire large data set has not been typically feasible, because the data needs to be imported into R first. 

Calculating bins of the *mpg* variable would involve the following formula:

```r
(3 * ifelse(
  as.integer(floor(((mpg) - min(mpg, na.rm = TRUE))/3)) == 
    as.integer((max(mpg, na.rm = TRUE) - min(mpg, na.rm = TRUE))/3), 
  as.integer(floor(((mpg) - min(mpg, na.rm = TRUE))/3)) - 1,
  as.integer(floor(((mpg) - min(mpg, na.rm = TRUE))/3)))
) + min(mpg, na.rm = TRUE)
```

The formula breaks down the creation of bins 3 mpg wide into a combination of the most basic aggregation functions.

The formula can run easily inside the `mutate()` command.  It creates as many discrete bins size 3 mpg, and passes the minimum value as the "label" of that bin.


```{r}
mtcars %>%
  mutate(
    bins =
      (3 * ifelse(
        as.integer(floor(((mpg) - min(mpg, na.rm = TRUE))/3)) == 
          as.integer((max(mpg, na.rm = TRUE) - min(mpg, na.rm = TRUE))/3), 
        as.integer(floor(((mpg) - min(mpg, na.rm = TRUE))/3)) - 1,
        as.integer(floor(((mpg) - min(mpg, na.rm = TRUE))/3)))
      ) + min(mpg, na.rm = TRUE)
  ) %>%
  select(mpg, bins) %>%
  head()
```

```
##    mpg bins
## 1 21.0 19.4
## 2 21.0 19.4
## 3 22.8 22.4
## 4 21.4 19.4
## 5 18.7 16.4
## 6 18.1 16.4
```

The same R formula can run inside Spark, `dplyr` will translate the R code into a SQL statement.  The results can then be collected into R for visualizing. 

```{r}
bins <- cars %>%
  mutate(
    bins =
      (3 * ifelse(
        as.integer(floor(((mpg) - min(mpg, na.rm = TRUE))/3)) == 
          as.integer((max(mpg, na.rm = TRUE) - min(mpg, na.rm = TRUE))/3), 
        as.integer(floor(((mpg) - min(mpg, na.rm = TRUE))/3)) - 1,
        as.integer(floor(((mpg) - min(mpg, na.rm = TRUE))/3)))
      ) + min(mpg, na.rm = TRUE)
  ) %>%
  count(bins) %>%
  collect()

bins
```

```
## # A tibble: 7 x 2
##    bins     n
##   <dbl> <dbl>
## 1  19.4     6
## 2  22.4     3
## 3  13.4     8
## 4  10.4     3
## 5  16.4     6
## 6  28.4     4
## 7  25.4     2
```

Since the bins and counts have been pre-calculated, a simple column plot is used.  

```{r}
bins %>%
  ggplot() +
  geom_col(aes(bins, n))
```

```{r, eval = TRUE, fig.width=10, fig.height=5, fig.cap='Plot from Spark', fig.align = 'center', echo = FALSE}
render_image("images/analysis-visualizations-2.png")
```


#### Using dbplot

The `dbplot` package provides helper function for plotting with remote data.  The pakcage uses `dplyr` to push the calculations to Spark and collects the results.

The `dbplot_histogram()` function will have Spark calculate the bins and the count per bin, and then outputs a `ggplot` object.  It accepts a `binwidth` argument.

```{r}
library(dbplot)

cars %>%
  dbplot_histogram(mpg, binwidth = 3)
```
```{r, eval = TRUE, fig.width=10, fig.height=5, fig.cap='Plot from Spark', fig.align = 'center', echo = FALSE}
render_image("images/analysis-visualizations-3.png")
```

Because it is a `ggplot` object, it can be further refined if needed.

```{r}
cars %>%
  dbplot_histogram(mpg, binwidth = 3) +
  labs(title = "Histogram of Miles Per Galon") +
  theme_bw()
```
```{r, eval = TRUE, fig.width=10, fig.height=5, fig.cap='Plot from Spark', fig.align = 'center', echo = FALSE}
render_image("images/analysis-visualizations-4.png")
```
The package also provides a way to obtain the raw results via the `db_compute_bins()` package.

```{r}
cars %>%
  db_compute_bins(mpg, binwidth = 3) 
```

```
## # A tibble: 7 x 2
##     mpg count
##   <dbl> <dbl>
## 1  19.4     6
## 2  22.4     3
## 3  13.4     8
## 4  10.4     3
## 5  16.4     6
## 6  28.4     4
## 7  25.4     2
```

### Scatter plots

Scatter plots are incredibly useful because of their multiple applications.  Very commonly, they are used to compare the relationship between two contionus variables.  The problems that arise when trying to use this visualization with a large amount of data are:

- Performance problems - Too many single dots have to be calculated and drawn.

- Perception problems - It becomes hard to see the true amount of dots in a single area.

No amount of "pushing the computation" to Spark will help with this problem becuase the data has to be plotted in individual dots.  

The best alternative is to find a plot type that represents the x/y relationship and concentration in a way that it is easy to percieve and to "physically" plot.  The *raster* plot may be the best answer.  It returns a grid of x/y positions and the results of a given aggregation usually represented by the collor of the square.  

The `dbplot` package provides funcitonality that helps with this kind of plotting.  The first one is `dbplot_raster()`.  It is the way to quickly run a visualization of this kind.

```{r}
cars %>%
  dbplot_raster(mpg, wt, resolution = 5)
```
```{r, eval = TRUE, fig.width=10, fig.height=5, fig.cap='Plot from Spark', fig.align = 'center', echo = FALSE}
render_image("images/analysis-visualizations-5.png")
```
As shown in *figure []*, the plot returns a grid no bigger than 5x5.  This limits the number of records that need to be collected into R to 25.  

If the user prefers to either visualize, or simply obtain the result data, the `db_compute_raster()` and `db_compute_raster2()` functions return the data.  The `db_compute_raster2()` includes the upper and lower bounds of each square.

```{r}
cars %>%
  db_compute_raster2(mpg, wt, resolution = 5)
```
```
## # A tibble: 9 x 5
##     mpg    wt `n()` mpg_2  wt_2
##   <dbl> <dbl> <dbl> <dbl> <dbl>
## 1  19.8  2.30     5  24.5  3.08
## 2  19.8  3.08     3  24.5  3.86
## 3  15.1  3.08    10  19.8  3.86
## 4  10.4  3.08     3  15.1  3.86
## 5  15.1  3.86     1  19.8  4.64
## 6  10.4  4.64     3  15.1  5.42
## 7  29.2  1.51     4  33.9  2.30
## 8  24.5  1.51     2  29.2  2.30
## 9  15.1  2.30     1  19.8  3.08
```

## Model


## Communicate

