```{r include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
source("r/render.R")
source("r/plots.R")
library(ggplot2)
```

# Extensions {#extensions}

> "I try to know as many people as I can. You never know which one you’ll need."
>
> --- Tyrion Lannister

In the previous chapter, [Tuning](tuning), you learned how Spark processes data at large-scale by allowing users to configure the cluster resources, partition data implicitly or explicitly, execute commands across distributed compute nodes, shuffle data across them when needed, cache data to improve performance and serialize data efficiently over the network. You also learned how to configure the different Spark settings used while connecting, submitting a job, running and application and particular settings applicable only to R and R extensions that we will present in this chapter.

The [Analysis](#analysis), [Modeling](#modeling) and Data chapters provided a foundation to read and understand most datasets. However, the functionality that was presented was scoped to Spark's built-in features and tabular datasets. This chapter will go beyond tabular data and explore how to analyze and model networks of interconnected objects through graph processing, read genomics datasets, prepare data for deep learning, analyze geographic datasets and use advanced modeling libraries like H2O and XGBoost over large-scale datasets.

The combination of all the content presented in all the previous chapters should take care of most of your large-scale computing needs. However, for those few use cases where functionality is still lacking, the following chapters will teach you provide the tools to extend Spark yourself; either, through custom R transformation, custom Scala code or through recent new execution mode in Spark that enable analyzing real-time datasets. Although, before reinventing the wheel, we will present all the extensions available when using Spark with R.

## Overview

In the Introduction chapter we presented the R community as a vibrant group of individuals collaborating with each other in many ways, one of them, by moving open science forward by creating R packages that can be installed from CRAN. In a similar way, but in a much smaller scale, the R community has contributed extensions that increase the functionality initially supported in Spark and R. Spark itself also provides support for creating Spark extensions and, in-fact, many R extensions make use of Spark extensions.

Extensions are constantly being created so this section will be outdated at any given point in time, in addition, we might not be even aware of many Spark and R extensions; however, at the very least we can track the extensions that are available in CRAN by looking at the "reverse imports" for `sparklyr` in CRAN. [@extensions-sparklyr-cran] Extensions and R packages published in CRAN tend to be the most stable since when a package is published in CRAN, it will go through a review process which increases the overall quality of a contribution.

While we wish we could present all the extensions, we will present next the extensions that should be the most interesting to most readers. You can find additional ones under the [github.com/r-spark](https://github.com/r-spark) organization or by searching repos in GitHub with the `sparklyr` tag.

rsparkling
: The `rsparkling` extensions allows you to use H2O and Spark from R. This extension is what we would consider advanced modeling in Spark. While Spark's built-in modeling library, Spark MLlib, is quite useful in many cases; H2O's modeling capabilities can compute additional statistical metrics and can proivide performance and scalability improvements over Spark MLlib. We, ourselves, have not performed detailed comparisons nor benchamarks between MLlib and H2O; so this is something you will have to research on your own to create a complete picture of when to use H2O's capabilities.

graphframes
: The `graphframes` extensions adds support to process graphs in Spark. A graph is a structure that describes a set of objects in which some pairs of the objects are in some sense related. As you learned in the Introduction chapter, ranking web pages was an early motivation to develop precursos to Spark powered by MapReduce; web pages happen to form a graph if you consider a link between pages as the relationship between each pair of pages. Computing operations likes PageRank over graphs can be quite useful in web search and social networks to mention a few applications.

sparktf
: The `sparktf` extension provides support to write TensorFlow records in Spark. TensorFlow is one of the leading deep learning frameworks and it is often used with large amounts of numerical data represented as TensorFlow records, a file format optimized for TensorFlow. Spark it is often used to process unstructured and large-scale datasets into smaller numerical datasets that can easily fit into a GPU. You can use this extension to save datasets in the TensorFLow record file format.

xgboost
: The `xgboost` extension brings the well-known XGBoost modeling library to the world of large-scale computing. XGBoost is a scalable, portable and distributed library for gradient boosting. It became well known in the machine learning competition circles after its use in the winning solution of the Higgs Machine Learning Challenge [@extensions-higgs-challenge] and has remain popular in other Kaggle competitions since then.

variantspark
: The `variantspark` extension provides an interface to use Variant Spark, a scalable toolkit for genome-wide association studies (GWAS). It currently provides functionality to build random forest models, estimating variable importance and reading Variant Call Format (VCF) files. While there are other random forest implementations in Spark, most of them are not optimized to deal with GWAS datasets, which usually come with thousands of samples and millions of variables.

geospark
: The `geospark` extensions enables us to load and query large-scale geographic datasets. Usually datasets containing latitude and longitude points or complex areas defined in the Well-known Text (WKT) format, a text markup language for representing vector geometry objects on a map.

Before you learn how and when to use each extension, we should first briefly explain how extensions can be used with R and Spark.

First, an Spark extension is just and R package that happens to be aware of Spark. As any other R package, you will first have to install the R package. Once installed, it is important to know that you will need to reconnect to Spark before the extension can be used. So, in general, the pattern you should follow goes as follows:

```{r}
library(sparkextension)
library(sparklyr)

sc <- spark_connect(master = "<master>")
```

Notice that `sparklyr` is loaded after the extensions to allow the extension to register properly. If you had to install and load a new extension you would simply have to disconnect first using `spark_disconnect(sc)` and repeat the steps above with the new extension.

It's not hard to install and use Spark extensions from R; however, each extension can be a world on it's own so most of the time you will have to spend time understand what the extension is, when to use it and how to use it properly. The first extension you will learn about is the `rsparkling` extension which enables you to use H2O in Spark with R.

## H2O

[H2O](https://www.h2o.ai/) is open-source software for large-scale modeling created by H2O.ai, which allows you to fit thousands of potential models as part of discovering patterns in data. You can consider using H2O to complement or replace Spark's default modeling algorithms. It is common to Spark's default modeling algorithms and transition to H2O when Spark's algorithms fall short or when advanced functionality (like additional modeling metrics or automatic model selection) are desired

We can't do justice to H2O's great modeling capabilities in a single paragraph, explaining H2O properly will require a book in itself. Instead, we would like to recommend reading the "Practical machine learning with H2O" [@extensions-practical-h2o] book to explore in-depth H2O's modeling algorithms and features. In the meantime, you can use this section as a brief guide to get started using H2O in Spark with R.

In order to use H2O with Spark, it is important to know that there are four compoinents involved: H2O, Sparkling Water, [rsparkling][rsparkling](https://github.com/h2oai/sparkling-water/tree/master/r) and Spark. Sparkling Water allows users to combine the fast, scalable machine learning algorithms of H2O with the capabilities of Spark. You can think of Sparkling Water as a component bridging Spark with H2O and `rsparkling` as the R front-end for Sparkling Water, this is illustrated in Figure \@ref(fig:extensions-h2o-diagram).

```{r extensions-h2o-diagram, eval=TRUE, echo=FALSE, fig.height=3, fig.width=8, fig.align='center', fig.cap='H2O components with Spark and R'}
render_nomnoml("
#spacing: 20
#padding: 16
[R | 
  [rsparkling]
  [sparklyr]
]->[Spark |
  [Sparkling Water]
  [H2O]
]
", "images/extensions-h2o-diagram.png")
```

First, install `rsparkling` and `h2o`:

```{r}
install.packages("rsparkling")
install.packages("h2o")
```

Is is then important to notice that you need to use compatible versions of Spark, Sparkling Water and H2O. So let's start by checking the version of H2O by running,

```{r eval=TRUE}
packageVersion("h2o")
```

Then we can explore the Spark and Sparkling Water versions that are compatible with `h2o_release_table()`.

```{r extensions-h2o-versions}
rsparkling::h2o_release_table() %>%
  dplyr::filter(H2O_Version == !!as.character(packageVersion("h2o")))
```
```
# A tibble: 4 x 5
  Spark_Version Sparkling_Water_V… H2O_Version H2O_Release_Name H2O_Release_Patch…
          <dbl> <fct>              <fct>       <fct>                         <int>
1           2.4 2.4.5              3.22.1.3    rel-xu                            3
2           2.3 2.3.23             3.22.1.3    rel-xu                            3
3           2.2 2.2.34             3.22.1.3    rel-xu                            3
4           2.1 2.1.48             3.22.1.3    rel-xu                            3
```

We can then connect with the supported Spark versions as follows, you will have to adjust the `master` parameter for your particular cluster.

```{r}
library(rsparkling)
library(sparklyr)
library(h2o)

sc <- spark_connect(master = "local", version = "2.3")
cars <- copy_to(sc, mtcars)
```

H2O provides a web interface which can help you monitor training and access much of H2O's functionality. The web interface can be accessed through `h2o_flow(sc)`, it is reffered to as H2O Flow and is shown in Figure \@ref(fig:extensions-h2o-flow).

```{r extensions-h2o-flow, eval=TRUE, fig.width=4, fig.align='center', echo=FALSE, fig.cap='H2O Flow Interface using Spark with R'}
render_image("images/extensions-h2o-flow.png")
```

When using H2O, you will have to convert your Spark DataFrame into and H2O DataFrame through `as_h2o_frame`:

```{r}
cars_h2o <- as_h2o_frame(sc, cars)
cars_h2o
```
```
   mpg cyl disp  hp drat    wt  qsec vs am gear carb
1 21.0   6  160 110 3.90 2.620 16.46  0  1    4    4
2 21.0   6  160 110 3.90 2.875 17.02  0  1    4    4
3 22.8   4  108  93 3.85 2.320 18.61  1  1    4    1
4 21.4   6  258 110 3.08 3.215 19.44  1  0    3    1
5 18.7   8  360 175 3.15 3.440 17.02  0  0    3    2
6 18.1   6  225 105 2.76 3.460 20.22  1  0    3    1

[32 rows x 11 columns] 
```

Then you can use many of the modeling functions available in the `h2o` package with ease. For instance, we can fit a generalized linear model with ease:

```{r eval=FALSE}
model <- h2o.glm(x = c("wt", "cyl"),
                 y = "mpg",
                 training_frame = cars_h2o,
                 lambda_search = TRUE)
```
```{r extensions-rsparkling, echo=FALSE}
saveRDS(model, "data/09-extensions-h2o-glm.rds")
```

H2O provides additional metrics not necessarily available in Spark's modeling algorithms, the model that we just fit  `Residual Deviance` is provided in the model while this would not be a standard metric when using Spark MLlib.

```{r}
model
```
```
...
MSE:  6.017684
RMSE:  2.453097
MAE:  1.940985
RMSLE:  0.1114801
Mean Residual Deviance :  6.017684
R^2 :  0.8289895
Null Deviance :1126.047
Null D.o.F. :31
Residual Deviance :192.5659
Residual D.o.F. :29
AIC :156.2425
```

Then you can run prediction over the generalized linear model model, a similar approach would work for many other models available in H2O:

```{r}
predictions <- as_h2o_frame(sc, copy_to(sc, data.frame(wt = 2, cyl = 6)))
h2o.predict(model, predictions)
```
```
   predict
1 24.05984

[1 row x 1 column]
```

H2O can also be used to perform automatic training and tuning of many models; meaning that, H2O can choose which model to use for you using `h2o.automl`:

```{r}
automl <- h2o.automl(x = c("wt", "cyl"), y = "mpg",
                     training_frame = cars_h2o,
                     max_models = 20,
                     seed = 1)
```

For this particular dataset, H2O finds out that XGBoost is a better fit than GLM. Specifically, the H2O explored using XGBoost, Deep Learning, GLM and a Stacked Ensemble.

```{r}
automl
```
```
model_id              mean_residual_deviance     rmse      mse      mae     rmsle
1 XGBoost_1_...       6.627278               2.574350 6.627278 2.066412 0.1329469
2 DeepLearning_...    6.945850               2.635498 6.945850 2.209201 0.1258008
3 XGBoost_grid_1_...  7.025614               2.650587 7.025614 2.192791 0.1339280
4 XGBoost_grid_1_...  7.266691               2.695680 7.266691 2.167930 0.1331849
5 GLM_grid_...        7.416367               2.723301 7.416367 2.184664 0.1269808
6 StackedEnsemble...  7.596133               2.756108 7.596133 2.029900 0.1340302
```

Many additional examples are available under [spark.rstudio.com/guides/h2o](http://spark.rstudio.com/guides/h2o/), you can also request help from [github.com/h2oai/sparkling-water/tree/master/r](https://github.com/h2oai/sparkling-water/tree/master/r), the official GitHub repository for the `rsparkling` package.

The next extension, `graphframes`, will allow you to process large-scale relational datasets; however, before you start using it, make make sure to disconnect with `spark_disconnect(sc)` since using a different extensions requires you to reconnect to Spark.

## Graphs

The first paper in the history of graph theory was written by Leonhard Euler on the Seven Bridges of Königsberg in 1736. The problem was to devise a walk through the city that would cross each of those bridges once and only, the original diagram is shown in Figure \@ref(fig:extensions-eulers-paths).

```{r extensions-eulers-paths, eval=TRUE, fig.width=4, fig.align='center', echo=FALSE, fig.cap='Seven Bridges of Königsberg from the Euler Archive'}
render_image("images/extensions-eulers-paths.png")
```

Today, a graph is defined as an ordered pair $G=(V,E)$, with $V$ a set of vertices (nodes or points) and $E \subseteq \{\{x, y\} | (x, y) ∈ \mathrm{V}^2 \land x \ne y\}$ a set of edges (links or lines) which are either an unordered pair for *undirected graphs* or an ordered pair for *directed graphs*. The former describing links where the direction does not matter and the latter linked where it does.

As a simple example, we can use the `highschool` dataset from the `ggraph` package which tracks friendship among high school boys. In this dataset, the vertices are the students and the edges describe pairs of students who happen to be friends in a particular year.

```{r}
library(ggraph)
library(igraph)

highschool
```
```
# A tibble: 506 x 3
    from    to  year
   <dbl> <dbl> <dbl>
 1     1    14  1957
 2     1    15  1957
 3     1    21  1957
 4     1    54  1957
 5     1    55  1957
 6     2    21  1957
 7     2    22  1957
 8     3     9  1957
 9     3    15  1957
10     4     5  1957
# … with 496 more rows
```

While the highschool dataset can be easily processed in R, even medium size graph datasets can be hard to process without distributing this work across a cluster of machines which Spark is suited for. Spark supports processing graphs through the [graphframes](https://github.com/rstudio/graphframes) extension which in turn uses the [GraphX](https://spark.apache.org/graphx/) Spark component. GraphX is Apache Spark's API for graphs and graph-parallel computation and allows you to seamlessly work with both graphs and collection, it's comparable in performance to the fastest specialized graph processing systems and provides a growing library of graph algorithms.

A graph in Spark is also represented as a dataframe of edges and vertices; however, the format is slightly different since we will need to construct a dataframe for vertices. Lets first install the [GraphFrames](https://graphframes.github.io/) extension,

```{r}
install.packages(graphframes)
```

Followed by connecting, copying the `highschool` dataset and transforming the graph to the format that this extension expects, we will scope this dataset to the friendships of year 1957.

```{r extensions-graphframes}
library(graphframes)
library(sparklyr)
library(dplyr)

sc <- spark_connect(master = "local", version = "2.1.0")
highschool_tbl <- copy_to(sc, highschool, "highschool") %>%
  filter(year == 1957) %>%
  transmute(from = as.character(as.integer(from)),
            to = as.character(as.integer(to)))

from_tbl <- highschool_tbl %>% distinct(from) %>% transmute(id = from)
to_tbl <- highschool_tbl %>% distinct(to) %>% transmute(id = to)

vertices_tbl <- distinct(sdf_bind_rows(from_tbl, to_tbl))
edges_tbl <- highschool_tbl %>% transmute(src = from, dst = to)
```

The `vertices_tbl` table is expected to have a single `id` column:

```
# Source: spark<?> [?? x 1]
   id   
   <chr>
 1 1    
 2 34   
 3 37   
 4 43   
 5 44   
 6 45   
 7 56   
 8 57   
 9 65   
10 71   
# … with more rows
```

While the `edges_tbl` is expected to have a `src` and `dst` columns:

```
# Source: spark<?> [?? x 2]
   src   dst  
   <chr> <chr>
 1 1     14   
 2 1     15   
 3 1     21   
 4 1     54   
 5 1     55   
 6 2     21   
 7 2     22   
 8 3     9    
 9 3     15   
10 4     5    
# … with more rows
```

You can now create a GraphFrame,

```{r}
graph <- gf_graphframe(vertices_tbl, edges_tbl)
```

We can now use this graph to start analyzing this dataset. For instance, by finding out how many friends on average every one has, this is reffered as the degree or valency of a vertex:

```{r}
gf_degrees(graph) %>% summarise(friends = mean(degree))
```
```
# Source: spark<?> [?? x 1]
  friends
    <dbl>
1    6.94
```

We can then find what the shortest path to some specific vertex (person for this dataset). Since the data is annonimized, we can just pick the person identified as $33$ and find how many degrees of separation exist between them:

```{r}
gf_shortest_paths(graph, 33) %>%
  filter(size(distances) > 0) %>%
  mutate(distance = explode(map_values(distances))) %>%
  select(id, distance)
```
```
# Source: spark<?> [?? x 2]
   id    distance
   <chr>    <int>
 1 19           5
 2 5            4
 3 27           6
 4 4            4
 5 11           6
 6 23           4
 7 36           1
 8 26           2
 9 33           0
10 18           5
# … with more rows
```

Finally, we can compute PageRank over this graph, which is named after Google's foudner Larry Page: 

```{r echo=FALSE}
model <- gf_graphframe(vertices_tbl, edges_tbl) %>%
  gf_pagerank(reset_prob = 0.15, max_iter = 10L)

highschool_tbl %>% collect() %>%
  saveRDS("data/09-extensions-graphframes-highschool.rds")
```
```{r extensions-graphframes-code}
gf_graphframe(vertices_tbl, edges_tbl) %>%
  gf_pagerank(reset_prob = 0.15, max_iter = 10L)
```
```
GraphFrame
Vertices:
  Database: spark_connection
  $ id       <dbl> 12, 12, 14, 14, 27, 27, 55, 55, 64, 64, 41, 41, 47, 47, 6…
  $ pagerank <dbl> 0.3573460, 0.3573460, 0.3893665, 0.3893665, 0.2362396, 0.…
Edges:
  Database: spark_connection
  $ src    <dbl> 7, 7, 7, 7, 7, 7, 7, 7, 7, 7, 7, 7, 7, 7, 7, 7, 12, 12, 12,…
  $ dst    <dbl> 17, 17, 17, 17, 17, 17, 17, 17, 17, 17, 17, 17, 17, 17, 17,…
  $ weight <dbl> 0.25000000, 0.25000000, 0.25000000, 0.25000000, 0.25000000,…
```

To give you some insights into this dataset, Figure \@ref(fig:extensions-graph-pagerank) plots this chart using the `ggraph` and highlights the highest PageRank scores fot this dataset, 

```{r extensions-graph-pagerank-create, eval=FALSE, echo=FALSE}
library(ggraph)
library(igraph)
highschool_rdf <- readRDS("data/09-extensions-graphframes-highschool.rds")
highschool_rdf %>% graph_from_data_frame(directed = FALSE) %>%
  ggraph(layout = 'kk') + 
    geom_edge_link(alpha = 0.2,
                   arrow = arrow(length = unit(2, 'mm')),
                   end_cap = circle(2, 'mm'),
                   start_cap = circle(2, 'mm')) + 
    geom_node_point(size = 2, alpha = 0.4) + theme_light() +
    annotate("point", x = -1.18, y = -3.55, size = 3) +
    annotate("point", x = 6.25, y = 2.85, size = 3) + xlab("") + ylab("") +
    ggsave("images/extensions-graph-pagerank.png", width = 10, height = 5)
```
```{r extensions-graph-pagerank, eval=TRUE, fig.width=4, fig.align='center', echo=FALSE, fig.cap='Highschool ggraph dataset with highest pagerank highlighted'}
render_image("images/extensions-graph-pagerank.png")
```

There are many more graph algorithms provided in `graphframes`, to mention some: bread depth search, connected components, label propagation for detecting communities, strongly connected components and triangle count. For questions on this extension reffer to the official GitHub repo, [github.com/rstudio/graphframes](https://github.com/rstudio/graphframes). We will now present a popular gradient boosting framework.

## XGBoost

A decision tree is a flowchart-like structure in which each internal node represents a test on an attribute, each branch represents the outcome of the test, and each leaf node represents a class label. For example, the diagram in Figure \@ref(fig:extensions-decision-diagram) shows an a decision tree that could help classify if an employee is likely to leave given a set of factors, like job satisfaction and overtime. When a decision tree is used to predict continuous variables instead of discrete outcomes, say, how likely someone is to leave a company,  decision trees are reffered to as regression trees.

```{r extensions-decision-diagram, eval=TRUE, echo=FALSE, fig.height=3, fig.width=8, fig.align='center', fig.cap='A Decision tree to predict job attrition based on known factors'}
render_nomnoml("
[0.5|Job Satisfaction > 0.2?
]-Yes[0.3|Over Time Hours > 10?]
[0.5]-No[0.7|Job Level > 2?]
[0.3]-Yes[0.65]
[0.3]-No[0.20]
[0.7]-Yes[0.15]
[0.7]-No[0.70]
", "images/extensions-decision-diagram.png")
```

While a decision tree representation is quite easy to understand and to interpret, finding out the decisions in the treee requires mathematical techniquest like gradient descent to find a local minimum. Gradient descent takes steps proportional to the negative of the gradient of the function at the current point. The gradient is represented by $\nabla$, the learning rate by $\gamma$ and one simply starts from a given state $a_n$ and compute the next iteration $a_{n+1}$ by simply following the direction of the gradient:

$$
a_{n+1} = a_n - \gamma \nabla F(a_n)
$$

XGBoost is an open-source software library which provides a gradient boosting framework. It aims to provide a scalable, portable and distributed gradient boosting for training gradient-boosted decision trees (GBDT) and gradient-boosted regression trees (GBDT). Gradient-boosted means xgboost uses gradient descent and boostrin, which is a technique that chooses each predictor sequentially.

`sparkxgb` is an extension that you can use to train XGBoost models in Spark. To use this extension, first install it from CRAN:

```{r}
install.packages("sparkxgb")
```

Then you would need to import the `sparkxgb` extension followed by your usual Spark connection code, adjusting `master` as needed:

```{r}
library(sparkxgb)
library(sparklyr)
library(dplyr)

sc <- spark_connect(master = "local")
```

For this example, we will use the `attrition` dataset from the `rsample` package which you would need to install with `install.packages("rsample")`. This dataset is a fictional dataset created by IBM data scientists to uncover the factors that lead to employee attrition.

```{r}
attrition <- copy_to(sc, rsample::attrition)
attrition
```
```
# Source: spark<?> [?? x 31]
     Age Attrition BusinessTravel DailyRate Department DistanceFromHome
   <int> <chr>     <chr>              <int> <chr>                 <int>
 1    41 Yes       Travel_Rarely       1102 Sales                     1
 2    49 No        Travel_Freque…       279 Research_…                8
 3    37 Yes       Travel_Rarely       1373 Research_…                2
 4    33 No        Travel_Freque…      1392 Research_…                3
 5    27 No        Travel_Rarely        591 Research_…                2
 6    32 No        Travel_Freque…      1005 Research_…                2
 7    59 No        Travel_Rarely       1324 Research_…                3
 8    30 No        Travel_Rarely       1358 Research_…               24
 9    38 No        Travel_Freque…       216 Research_…               23
10    36 No        Travel_Rarely       1299 Research_…               27
# … with more rows, and 25 more variables: Education <chr>,
#   EducationField <chr>, EnvironmentSatisfaction <chr>, Gender <chr>,
#   HourlyRate <int>, JobInvolvement <chr>, JobLevel <int>, JobRole <chr>,
#   JobSatisfaction <chr>, MaritalStatus <chr>, MonthlyIncome <int>,
#   MonthlyRate <int>, NumCompaniesWorked <int>, OverTime <chr>,
#   PercentSalaryHike <int>, PerformanceRating <chr>,
#   RelationshipSatisfaction <chr>, StockOptionLevel <int>,
#   TotalWorkingYears <int>, TrainingTimesLastYear <int>,
#   WorkLifeBalance <chr>, YearsAtCompany <int>, YearsInCurrentRole <int>,
#   YearsSinceLastPromotion <int>, YearsWithCurrManager <int>
```

To build an XGBoost model in Spark use `xgboost_classifier()`, we will compute attrition against all other features by using the `Attrition ~ .` formula and specify two for the number of classes since the attrition attribute tracks only whether an employee leaves or not. Then you can use `ml_predict()` to predict over large-scale datasets:

```{r}
xgb_model <- xgboost_classifier(attrition,
                                Attrition ~ .,
                                num_class = 2,
                                num_round = 50,
                                max_depth = 4)

xgb_model %>%
  ml_predict(attrition) %>%
  select(Attrition, predicted_label, starts_with("probability_")) %>%
  glimpse()
```
```
Observations: ??
Variables: 4
Database: spark_connection
$ Attrition       <chr> "Yes", "No", "Yes", "No", "No", "No", "No", "No", "No", …
$ predicted_label <chr> "No", "Yes", "No", "Yes", "Yes", "Yes", "Yes", "Yes", "Y…
$ probability_No  <dbl> 0.753938094, 0.024780750, 0.915146366, 0.143568754, 0.07…
$ probability_Yes <dbl> 0.24606191, 0.97521925, 0.08485363, 0.85643125, 0.927375…
```

XGBoost became well known in the competition circles after its use in the winning solution of the Higgs Machine Learning Challenge which uses the ATLAS experiment to identify the Higgs boson. Since then, it has become a popular model and used for a large number of Kaggle competitions. However, decision trees could prove limiting specially in datasets with non tabular data like images, audio and text which you can tackle with deep learning models.

## Deep Learning

A perceptron is a mathematical model introduced by Rosenblatt [@extensions-rosenblatt-perceptron] whom developed it as a theory for a hypothetical nervous system. The perceptron tries to mimic a neuron by mapping stimulae to numeric inputs that are weighted into an threshold function that activates only when enough stimulae is present, mathematically:

$$
f(x) =
\begin{cases}
1 & \sum_{i=1}^m w_i x_i + b > 0\\
0 & \text{otherwise}
\end{cases}
$$

Minsky found out that a single perceptron can only classify datasets that are linearly separable; however, he also presented in his perceptrons book [@extensions-minsky-perceptrons] that layering perceptrons would bring additional classification capabilities, the original diagram showcasing a multi-layered perceptron is presented in Figure \@ref(fig:extensions-minsky-layered).

```{r extensions-minsky-layered, eval=TRUE, echo=FALSE, fig.cap='Layered perceptrons as illustrated in the perceptrons book', fig.align = 'center'}
render_image("images/extensions-minsky-multi-layers.png")
```

Using Spark we can create a multi-layer perceptron classifier with `ml_multilayer_perceptron_classifier()` to classify and predict over large datasets. This technique was introduced by Hinton [@extensions-hinton-layered] and like XGBoost, it also makes use of gradient descent.

```{r}
nn_model <- ml_multilayer_perceptron_classifier(
  attrition,
  Attrition ~ Age + DailyRate + DistanceFromHome + MonthlyIncome,
  layers = c(4, 3, 2))

nn_model %>%
  ml_predict(attrition) %>%
  select(Attrition, predicted_label, starts_with("probability_")) %>%
  glimpse()
```
```
Observations: ??
Variables: 4
Database: spark_connection
$ Attrition       <chr> "Yes", "No", "Yes", "No", "No", "No", "No", "No", "No"…
$ predicted_label <chr> "No", "No", "No", "No", "No", "No", "No", "No", "No", …
$ probability_No  <dbl> 0.8439275, 0.8439275, 0.8439275, 0.8439275, 0.8439275,…
$ probability_Yes <dbl> 0.1560725, 0.1560725, 0.1560725, 0.1560725, 0.1560725,…
```

Notice that the columns must be numeric so you will have to manually convert them with feature transformin techniques presented in the Modeling chapter. It is natural to try to add more layers to classify more complex datasets; however, adding too many layers will cause the gradient to vanish and other techniques will have to use these deep layered networds also known as, deep learning models.

Deep learning models solve the vanishing gradient problem by making use of special activation functions, dropout, data augmentation and graphic processing units (GPUs). You can use Spark to retrieve and preprocess large datasets into numerical-only datasets that can fit in a GPU for deep learning training. Tensorflow is one of the most popular deep learning frameworks and supports a binary format known as TensorFlow Records.

You can write TensorFlow records using the `sparktf` in Spark which you can prepare to process in GPU instances with libraries like Keras or TensorFlow. `sparktf` is available on CRAN and can be installed as follows:

```{r}
install.packages("sparktf")
```

You can then preprocess large datasets in Spark and write it as TensorFlow records using `spark_write_tf()`:

```{r}
library(sparktf)
library(sparklyr)

sc <- spark_connect(master = "local")

copy_to(sc, iris) %>%
  ft_string_indexer_model(
    "Species", "label",
    labels = c("setosa", "versicolor", "virginica")
  ) %>%
  spark_write_tfrecord(path = "tfrecord")
```

Once trained, you can use the `tfdatasets` package to load the dataset followed by training with `keras` or `tensorflow`. To learn more about training deep learning models with Keras we recommend reading "Deep Learning with R". [@extensions-chollet-jj]

```{r}
tfdatasets::tfrecord_dataset("tfrecord/part-r-00000")
```
```
<DatasetV1Adapter shapes: (), types: tf.string>
```

Training deep learning models in a single local node with one or more GPUs is often enough for most applciations; however, recent state-of-the-art deep learning models train using distributed computing frameworks like Apache Spark. Distributed computing frameworks are used to achieve higher petaflops each day the systems spends training these models. OpenAI analyzed trends in AI and Compute [@extensions-openai-compute] which Figure \@ref(fig:extensions-distributed-training) adapted to show which systems were trained in local instances or distributed computing systems. It should be obvious from the figure that there is a trend in recent years to use distributed computing frameworks.

```{r echo=FALSE}
data.frame(
  system = c("AlexNet", "Dropout", "DQN", "Conv Nets", "GoogleNet", "Seq2Seq / VGG", "ResNets", "DeepSpeech2",
             "Xception", "TI7 Dota 1v1", "Neural Architecture Search", "Neural Machine Translation", "AlphaZero", "AlphaGo Zero"),
  petaflops = c(0.0058, 0.003, 0.0002, 0.009, 0.02, 0.1, 0.12, 0.30,
                4.5, 8.5, 40, 90, 500, 2000),
  year = as.Date(paste0(c("2012-02", "2012-06", "2013-09", "2013-08", "2013-08", "2013-07", "2015-11", "2015-11",
                          "2016-10", "2017-08", "2016-10", "2016-08", "2017-11", "2017-09"), "-01"), "%Y-%m-%d"),
  training = c("local", "local", "local", "local", "local", "local", "local", "local",
               "distributed", "distributed", "distributed", "distributed", "distributed", "distributed")) %>%
    ggplot(aes(x = year, y = petaflops, label = system, color = training)) +
      scale_color_grey(start = 0.2, end = 0.6) +
      labs(title = "Distributed Deep Learning",
           subtitle = "Training using distributed systems based on OpenAI analysis") +
      theme(legend.position = "right") +
      geom_point(size = 2.5) +
      geom_text(aes(label = system), hjust = -0.2, vjust = 0.4) +
      scale_y_log10(
        limits = c(1/10^4, 10^4),
        breaks = c(1/10^5, 1/10^4, 1/10^3, 1/10^2, 1/10^1, 1, 10^1, 10^2, 10^3, 10^4),
        labels = c("0.00001", "0.0001", "0.001", "0.01", "0.1", "1", "10", "100", "1000", "10000")
      ) +
      scale_x_date(name = "",
                   limits = as.Date(as.character(c(2011, 2020)), "%Y"),
                   breaks = as.Date(as.character(seq(2013, 2019, by = 2)), "%Y"),
                   labels = as.character(seq(2013, 2019, by = 2))) +
       ggsave("images/extensions-distributed-training.png", width = 10, height = 5)
```
```{r extensions-distributed-training, eval=TRUE, echo=FALSE, fig.cap='Training using distributed systems based on OpenAI analysis', fig.align = 'center'}
render_image("images/extensions-distributed-training.png")
```

Training large-scale deep learning models is possible in Spark and TensorFlow through frameworks like Horovod. Today, it's possible to use Horovod with Spark from R using the `reticualte` package since Horovod requires Python and Open MPI which goes beyond the scope of this book. Instead, we will introduce a different Spark extension in the domain of genomics.

## Genomics

The human genome consists of two copies of about three billion base pairs of DNA within the 23 chromosome pairs, Figure \@ref(fig:extensions-genomics-diagram) shows the organization of the genome into chromosomes. DNA strands are composed of nucleotides, each composed of one of four nitrogen-containing nucleobases: cytosine (C), guanine (G), adenine (A) or thymine (T). [@extensions-genomics-wikipedia] Since the DNA of all humans is nearly identical, we only need to store the differences from the reference genome in the form of a Variant Call Format (VCF) file.

```{r extensions-genomics-diagram, eval=TRUE, echo=FALSE, fig.cap='The idealized human diploid karyotype showing the organization of the genome into chromosomes', fig.align = 'center'}
render_image("images/extensions-genomics-diagram.png")
```

VariantSpark is a framework based on scala and spark to analyze genome datasets. It is being developed by CSIRO Bioinformatics team in Australia. VariantSpark was tested on datasets with 3000 samples each one containing 80 million features in either unsupervised clustering approaches and supervised applications, like classification and regression. VariantSpark can read VCF files and run analyses while using familiar Spark DataFrames. To get started, install `variantspark` from CRAN and connect to Spark:

```{r}
library(variantspark)
library(sparklyr)

sc <- spark_connect(master = "local")
vsc <- vs_connect(sc)
```

We can start by loading a VCF file,

```{r}
vsc_data <- system.file("extdata/", package = "variantspark")

hipster_vcf <- vs_read_vcf(vsc, file.path(vsc_data, "hipster.vcf.bz2"))
hipster_labels <- vs_read_csv(vsc, file.path(vsc_data, "hipster_labels.txt"))
labels <- vs_read_labels(vsc, file.path(vsc_data, "hipster_labels.txt"))
```

VariantSpark uses Random Forest to assign an Importance score to each tested variant reflecting its association to the interest phenotype. A variant with higher Importance score implies it is more strongly associated with the phenotype of interest. You can compute the Importance and transform it into a Spark table as follows,

```{r}
importance_tbl <- vs_importance_analysis(vsc, hipster_vcf, labels, n_trees = 100) %>%
  importance_tbl()

importance_tbl
```
```
# Source: spark<?> [?? x 2]
   variable    importance
   <chr>            <dbl>
 1 2_109511398 0         
 2 2_109511454 0         
 3 2_109511463 0.00000164
 4 2_109511467 0.00000309
 5 2_109511478 0         
 6 2_109511497 0         
 7 2_109511525 0         
 8 2_109511527 0         
 9 2_109511532 0         
10 2_109511579 0         
# … with more rows
```

You can then use `dplyr` and `ggplot2` to transform the output and visualize it,

```{r}
importance_df <- importance_tbl %>% 
  arrange(-importance) %>% 
  head(20) %>% 
  collect()

ggplot(importance_df) +
  aes(x = variable, y = importance) + 
  geom_bar(stat = 'identity') +          
  scale_x_discrete(limits = importance_df[order(importance_df$importance), 1]$variable) + 
  coord_flip()
```
```{r echo=FALSE}
ggplot(importance_df) +
  aes(x = variable, y = importance) + 
  labs(title = "Importance Analysis",
       subtitle = "Genomic importance analysis using variantspark") +
  scale_color_grey(start = 0.2, end = 0.6) +
  theme(legend.position = "right",
        axis.text.y = element_text(color = "#999999"),
        axis.text = ggplot2::element_text(size=8)) +
  geom_bar(stat = 'identity', fill = "#888888") +
  scale_x_discrete(limits = importance_df[order(importance_df$importance), 1]$variable) + 
  coord_flip() +
  ggsave("images/extensions-genomics-importance.png", width = 10, height = 5)
```
```{r extensions-genomics-importance, eval=TRUE, echo=FALSE, fig.cap='Genomic importance analysis using variantspark', fig.align = 'center'}
render_image("images/extensions-genomics-importance.png")
```

This concludes a brief introduction to genomic analysis in Spark using the variantspark extensions. Next, we will move away from micrososcopic genes, to macroscopic datasets that contain geographic locations across the world.

## Spatial 

[geospark](https://github.com/harryprince/geospark) enables distributed geospatial computing using a grammar compatible with [`dplyr`](https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf) and [`sf`](https://github.com/rstudio/cheatsheets/raw/master/sf.pdf) package which provides a set of tools for working with geospatial vectors.

You can install `geospark` from GitHub as follows:

```{r}
install.packages("remotes")
remotes::install_github("r-spark/geospark")
```

Then we will initialize the geospark extension and connect to Spark:

```{r}
library(geospark)
library(sparklyr)

sc <- spark_connect(master = "local")
register_gis(sc)
```

Next we will load a spatial dataset containing polygons and points.

```{r}
polygons <- system.file("examples/polygons.txt", package="geospark") %>%
  read.table(sep="|", col.names = c("area", "geom"))

points <- system.file("examples/points.txt", package="geospark") %>%
  read.table(sep = "|", col.names = c("city", "state", "geom"))

polygons_wkt <- copy_to(sc, polygons)
points_wkt <- copy_to(sc, points)
```

There are various spatial operations defined in `geospark`, which Figure \@ref(fig:extensions-geospark-operations) describes. These operations allow you to control how geospatial data should be queried based on overlap, intersection, disjoint sets, etc.

```{r extensions-geospark-operations, eval=TRUE, fig.width=4, fig.align='center', echo=FALSE, fig.cap='Spatial operations available in geospark.'}
render_image("images/extensions-geospark-operations.png")
```

For isntance, we can use these operations to find the polygons that contain a given set of points using `st_contains()`,

```{r}
library(dplyr)
polygons_wkt <- mutate(polygons_wkt, y = st_geomfromwkt(geom))
points_wkt <- mutate(points_wkt, x = st_geomfromwkt(geom))

sc_res <- inner_join(polygons_wkt,
                     points_wkt,
                     join = sql("st_contains(y,x)")) %>% 
  group_by(area, state) %>%
  summarise(cnt = n()) %>%
  head()
```
```
    # Source: spark<?> [?? x 3]
    # Groups: area
      area            state   cnt
      <chr>           <chr> <dbl>
    1 texas area      TX       10
    2 dakota area     SD        1
    3 dakota area     ND       10
    4 california area CA       10
    5 new york area   NY        9
```

You can also plot these datasets by collecting a subset of the entire dataset or aggregating the geometries in spark before collecting them, one package you should look into is the `sf` package. We will now start closing this chapter by presenting a couple troubleshooting techniques applicable to all extensions.

## Troubleshooting

When using a new extension for the first time, we recommend increasing the connection timeout since Spark will usually have to download extension dependencies and changing logging to verbose to help troubleshoot when the download process does not complete:

```{r}
config <- spark_config()
config$sparklyr.connect.timeout <- 3 * 60
config$sparklyr.log.console = TRUE

sc <- spark_connect(master = "local", config = config)
```

In addition, you should know that [Apache IVY](http://ant.apache.org/ivy/) is a popular dependency manager focusing on flexibility, simplicity, and is used by Apache Spark while installing extensions. When the connection fails while using an extension, consider clearing your [IVY Cache](http://ant.apache.org/ivy/history/2.0.0/settings/caches.html) by running:

```{r extensions-rsparkling-cache}
unlink("~/.ivy2/cache", recursive = TRUE)
```

In addition, you can also consider opening GitHub issues from the following extensions repos to get help from the extension authors:

- *rsparkling*: [github.com/h2oai/sparkling-water](https://github.com/h2oai/sparkling-water).
- *sparkxgb*: [github.com/rstudio/sparkxgb](https://github.com/rstudio/sparkxgb).
- *sparktf*: [github.com/rstudio/sparktf](https://github.com/rstudio/sparktf).
- *variantspark*: [github.com/r-spark/variantspark](https://github.com/r-spark/variantspark).
- *geospark*: [github.com/r-spark/geospark](https://github.com/r-spark/geospark).

## Recap

This chapter provideed a brief oveview on using some of the Spark extensions available in R, which happens to be as easy as installing a package. You then learned how to use the `rsparkling` extension which provides access to H2O in Spark to which provides additional modeling functionality like enhanced metrics and ability to automatically select models. We then jumped to `graphframes`, an extension to help you process relational datasets which are formaly referred as graphs; you learned how to compute simple connection metrics or run complex algorithms like pagerank.

THe XGBoost and Deep Learning sections provided alternate modeling techniques that use gradient descent, the former over decision trees and the latter over deep multi-layered perceptrons where Spark can be used to preprocess datasets into records that can then be later consument by TensorFlow and Keras using the `sparktf` extension. The last two sections untroduced extensions to process genomic and spatial datasets through the `variantspark` and `geospark` extensions.

These extensions and many more, provide a comprehensive library of advanced functionality that in combination with analysis and modeling techniques presented, should cover most tasks required to run in computing clusters. However, when functionality is lacking, you can consideer writting your own extension as we will present in the [Contributing](#contributing) or you can apply custom transformations over each partition using R code as we will described in the next chapter, [Ditributed R](#distributed-r).
