```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)

library(tidyverse)
```

# Modeling {#modeling}

While **this chatper has not been written**, a few resources and basic examples were made available to help out until this chapter is completed.

## Overview

MLlib is Apache Spark's scalable machine learning library and is available through `sparklyr`, mostly, with functions prefixed with `ml_`. The following table describes some of the modeling algorithms supported:

Algorithm | Function
----------|---------
Accelerated Failure Time Survival Regression | ml_aft_survival_regression()
Alternating Least Squares Factorization | ml_als()
Bisecting K-Means Clustering | ml_bisecting_kmeans()
Chi-square Hypothesis Testing | ml_chisquare_test()
Correlation Matrix | ml_corr()
Decision Trees | ml_decision_tree	()
Frequent Pattern Mining | ml_fpgrowth()
Gaussian Mixture Clustering | ml_gaussian_mixture()
Generalized Linear Regression | ml_generalized_linear_regression()
Gradient-Boosted Trees | ml_gradient_boosted_trees()
Isotonic Regression | ml_isotonic_regression()
K-Means Clustering | ml_kmeans()
Latent Dirichlet Allocation | ml_lda()
Linear Regression | ml_linear_regression()
Linear Support Vector Machines | ml_linear_svc()
Logistic Regression | ml_logistic_regression()
Multilayer Perceptron | ml_multilayer_perceptron()
Naive-Bayes | ml_naive_bayes()
One vs Rest | ml_one_vs_rest()
Principal Components Analysis | ml_pca()
Random Forests | ml_random_forest()
Survival Regression | ml_survival_regression()

To complement those algorithms, you will often also want to consider using the following feature transformers:

Transformer | Function
------------|---------
Binarizer | ft_binarizer()
Bucketizer | ft_bucketizer()
Chi-Squared Feature Selector | ft_chisq_selector()
Vocabulary from Document Collections | ft_count_vectorizer()
Discrete Cosine Transform  | ft_discrete_cosine_transform()
Transformation using dplyr | ft_dplyr_transformer()
Hadamard Product | ft_elementwise_product()
Feature Hasher | ft_feature_hasher()
Term Frequencies using Hashing | export(ft_hashing_tf)
Inverse Document Frequency | ft_idf()
Imputation for Missing Values | export(ft_imputer)
Index to String | ft_index_to_string()
Feature Interaction Transform | ft_interaction()
Rescale to [-1, 1] Range | ft_max_abs_scaler()
Rescale to [min, max] Range | ft_min_max_scaler()
Locality Sensitive Hashing | ft_minhash_lsh()
Converts to n-grams | ft_ngram()
Normalize using the given P-Norm | ft_normalizer()
One-Hot Encoding | ft_one_hot_encoder()
Feature Expansion in Polynomial Space | ft_polynomial_expansion()
Maps to Binned Categorical Features | ft_quantile_discretizer()
SQL Transformation | ft_sql_transformer()
Standardizes Features using Corrected STD | ft_standard_scaler()
Filters out Stop Words | ft_stop_words_remover()
Map to Label Indices | ft_string_indexer()
Splits by White Spaces | export(ft_tokenizer)
Combine Vectors to Row Vector | ft_vector_assembler()
Indexing Categorical Feature | ft_vector_indexer()
Subarray of the Original Feature | ft_vector_slicer()
Transform Word into Code | ft_word2vec()

## Supervised

Examples are reosurces are available in [spark.rstudio.com/mlib](http://spark.rstudio.com/mlib/).

## Unsupervised

### K-Means Clustering

Here is an example to get you started with K-Means:

```{r modeling-kmeans}
library(sparklyr)

# Connect to Spark in local mode
sc <- spark_connect(master = "local")

# Copy iris to Spark
iris_tbl <- sdf_copy_to(sc, iris, overwrite = TRUE)

# Run K-Means for Species using only Petal_Width and Petal_Length as features
iris_tbl %>%
  ml_kmeans(centers = 3, Species ~ Petal_Width + Petal_Length)
```

### Gaussian Mixture Clustering

Alternatevely, we can also cluster using [Gaussian Mixture Models](https://en.wikipedia.org/wiki/Mixture_model#Gaussian_mixture_model) (GMMs).

```{r modeling-gmm, echo=FALSE}
devtools::install_github("hadley/fueleconomy")

library(sparklyr)
sc <- spark_connect(master = "local", version = "2.3.0")
vehicles_tbl <- copy_to(sc, fueleconomy::vehicles, overwrite = TRUE)

predictions <- vehicles_tbl %>%
  ml_gaussian_mixture(~ hwy + cty, k = 3) %>%
  ml_predict() %>%
  collect()

saveRDS(predictions, "data/03-gaussian-mixture-prediction.rds")
```
```{r}
predictions <- copy_to(sc, fueleconomy::vehicles) %>%
  ml_gaussian_mixture(~ hwy + cty, k = 3) %>%
  ml_predict() %>% collect()

predictions %>%
  ggplot(aes(hwy, cty)) +
  geom_point(aes(hwy, cty, col = factor(prediction)), size = 2, alpha = 0.4) + 
  scale_color_discrete(name = "", labels = paste("Cluster", 1:3)) +
  labs(x = "Highway", y = "City") + theme_light()
```
```{r modeling-gmm-prediciton, eval=TRUE, fig.cap="Fuel economy data for 1984-2015 from the US EPA"}
predictions <- readRDS("data/03-gaussian-mixture-prediction.rds")
predictions %>%
  ggplot(aes(hwy, cty)) +
  geom_point(aes(hwy, cty, col = factor(prediction)), size = 2, alpha = 0.4) + 
  scale_color_discrete(name = "", labels = paste("Cluster", 1:3)) +
  labs(x = "Highway", y = "City") +
  theme_light()
```

## Broom

You can turn your `sparklyr` models into data frames using the `broom` package:

```{r modeling-broom, echo=FALSE}
library(sparklyr)
library(dplyr)
sc <- spark_connect(master = "local", version = "2.3.0")
cars_tbl <- spark_read_csv(sc, "cars", "input/")
model <- cars_tbl %>% ml_linear_regression(mpg ~ wt + cyl)
tidy_model <- list(
  broom::tidy(model),
  broom::glance(model),
  collect(broom::augment(model, cars_tbl))
)
saveRDS(tidy_model, "data/03-broom-examples.rds")
```
```{r modeling-broom-tidy, eval=TRUE, echo=FALSE}
tidy_model <- readRDS("data/03-broom-examples.rds")
```
```{r modeling-broom-lm}
model <- cars_tbl %>%
  ml_linear_regression(mpg ~ wt + cyl)

# Turn a model object into a data frame
broom::tidy(model)
```
```{r modeling-broom-model, echo=FALSE, eval=TRUE}
tidy_model[[1]]
```
```{r modeling-broom-glance}
# Construct a single row summary
broom::glance(model)
```
```{r modeling-broom-tidy2, echo=FALSE, eval=TRUE}
tidy_model[[2]]
```
```{r}
# Augments each observation in the dataset with the model
broom::augment(model, cars_tbl)
```
```{r echo=FALSE, rows.print=3, eval=TRUE}
tidy_model[[3]]
```

## Pipelines

Spark’s ML Pipelines provide a way to easily combine multiple transformations and algorithms into a single workflow, or pipeline.

Take a look at [spark.rstudio.com/guides/pipelines](http://spark.rstudio.com/guides/pipelines/) to learn about their purpose and functionality.
