\mainmatter

```{r include=FALSE, eval=TRUE}
knitr::opts_chunk$set(eval = FALSE)
source("r/render.R")
library(sparklyr)
library(ggplot2)
```

# Getting Started {#starting}

From R, getting started with Spark using `sparklyr` and a local cluster is as easy as running:

```{r starting-connect}
spark_install()
sc <- spark_connect(master = "local")
```

To make sure we can all run the code above and understand it, this section will walk you through the prerequisites, installing `sparklyr` and Spark, connecting to a local Spark cluster and briefly explaining how to use Spark.

However, if a Spark cluster and R environment have been made available to you, you do not need to install the [prerequisites](#starting-prerequisites) nor [install Spark](#starting-installing-spark) yourself. Instead, you should ask for the Spark `master` parameter and connect as follows; this parameter will be formally introduced under the [clusters](#clusters) and [connections](#connections) chapters.

```{r starting-connect-master}
sc <- spark_connect(master = "<cluster-master>")
```

## Prerequisites {#starting-prerequisites}

R can run in many platforms and environments; therfore, whether you use Windows, Mac or Linux, the first step is to install R from the [r-project.org](https://r-project.org/), detailed instructions are provided in the [Installing R](#appendix-install-r) appendix.

Most people use programming languages with tools to make them more productive; for R, RStudio would be such tool. Strictly speaking, RStudio is an Integrated Development Environment (or IDE), which also happens to support many platforms and environments. We strongly recommend you get RStudio installed if you haven't done so already, see details under the [Installing RStudio](#appendix-install-rstudio) appendix.

Additionally, since Spark is built in the Scala programming language which is run by the Java Virtual Machine, you also need to install Java 8 in your system. It is likely that your system already has Java installed, but you should still check the version and update if needed as described in the [Installing Java](#appendix-install-java) appendix.

## Installing sparklyr {#starting-install-sparklyr}

As many other R packages, `sparkylr` is available to be installed from [CRAN](https://CRAN.R-project.org/package=sparklyr) and can be easily installed as follows:

```{r starting-install-sparklyr, eval=FALSE}
install.packages("sparklyr")
```

The CRAN release of `sparklyr` contains the most stable version and it's the recommended version to use; however, to try out features being developed in `sparklyr`, you can install directly from GitHub using the `devtools` package. First, install the `devtools` package and then install `sparklyr` as follows:

```{r starting-install-sparklyr-devel, eval=FALSE}
install.packages("remotes")
remotes::install_github("rstudio/sparklyr")
```

The examples in this book assume you are using the latest version of `sparklyr`, you can verify your version is as new as the one we are using by running:

```{r}
packageVersion("sparklyr")
```
```
[1] ‘1.0.0’
```

## Installing Spark {#starting-installing-spark}

Start by loading `sparklyr`, 

```{r starting-install-spark-header, warning=FALSE, message=FALSE}
library(sparklyr)
```

This will makes all `sparklyr` functions available in R, which is really helpful; otherwise, we would have to run each `sparklyr` command prefixed with `sparklyr::`.

Spark can be easily installed by running `spark_install()`; this will install the latest version of Spark locally in your computer, go ahead and run `spark_install()`. Notice that this command requires internet connectivity to download Spark.

```{r starting-install-spark}
spark_install()
```

All the versions of Spark that are available for installation can be displayed by running:

```{r starting-install-available, eval=TRUE}
spark_available_versions()
```

A specific version can be installed using the Spark version and, optionally, by also specifying the Hadoop version. For instance, to install Spark 1.6.3, we would run:

```{r starting-install-install-version}
spark_install(version = "1.6")
```

You can also check which versions are installed by running:

```{r starting-install-installed}
spark_installed_versions()
```
```
  spark hadoop                              dir
7 2.3.1    2.7 /spark/spark-2.3.1-bin-hadoop2.7
```

The path where Spark is installed is referenced as Spark's home, which is defined in R code and system configuration settings with the `SPARK_HOME` identifier. When using a local Spark cluster installed with `sparklyr`, this path is already known and no additional configuration needs to take place.

Finally, in order to uninstall an specific version of Spark you can run `spark_uninstall()` by specifying the Spark and Hadoop versions, for instance:

```{r starting-install-uninstall}
spark_uninstall(version = "1.6.3", hadoop = "2.6")
```

**Note:** The default installation paths are `~/spark` for OS X and Linux and, `%LOCALAPPDATA%/spark` for Windows. To customize the installation path you can run `options(spark.install.dir = "<installation-path>")` before `spark_install()` and `spark_connect()`.

## Connecting to Spark {#starting-connect-to-spark}

It's important to mention that, so far, we've only installed a local Spark cluster. A local cluster is really helpful to get started, test code and troubleshoot with ease. Further chapters will explain where to find, install and connect to real Spark clusters with many machines, but for the first few chapters, we will focus on using local clusters.

To connect to this local cluster we simply run:

```{r starting-connect-local}
sc <- spark_connect(master = "local")
```

The `master` parameter identifies which is the "main" machine from the Spark cluster; this machine is often called the driver node. While working with real clusters using many machines, most machines will be worker machines and one will be the master. Since we only have a local cluster with only one machine, we will default to use `"local"` for now.

If connection fails, the [connections](#connections) chapter contains a [troubleshooting](#connections-troubleshooting) section which can help you resolve your connection issue.

## Using Spark {#starting-sparklyr-hello-world}

Now that you are connected, we can run a few simple commands. For instance, let's start by copying the `mtcars` dataset into Apache Spark using `copy_to()`.

```{r starting-copy-cars}
cars <- copy_to(sc, mtcars)
```

The data was copied into Spark but we can access it from R using the `cars` reference. To print it's contents we can simply type `cars`.

```{r starting-print-cars}
cars
```
```
# Source: spark<mtcars> [?? x 11]
     mpg   cyl  disp    hp  drat    wt  qsec    vs    am  gear  carb
   <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
 1  21       6  160    110  3.9   2.62  16.5     0     1     4     4
 2  21       6  160    110  3.9   2.88  17.0     0     1     4     4
 3  22.8     4  108     93  3.85  2.32  18.6     1     1     4     1
 4  21.4     6  258    110  3.08  3.22  19.4     1     0     3     1
 5  18.7     8  360    175  3.15  3.44  17.0     0     0     3     2
 6  18.1     6  225    105  2.76  3.46  20.2     1     0     3     1
 7  14.3     8  360    245  3.21  3.57  15.8     0     0     3     4
 8  24.4     4  147.    62  3.69  3.19  20       1     0     4     2
 9  22.8     4  141.    95  3.92  3.15  22.9     1     0     4     2
10  19.2     6  168.   123  3.92  3.44  18.3     1     0     4     4
# … with more rows
```

**Congrats!** You have successfully connected and loaded your first dataset into Spark.

Let's explain what's going on in `copy_to()`. The first parameter, `sc`, gives the function a reference to the active Spark Connection that was earlier created with `spark_connect()`. The second parameter specifies a dataset to load into Spark. Now, `copy_to()` returns a reference to the dataset in Spark which R automatically prints. Whenever a Spark dataset is printed, Spark will **collect** some of the records and display them for you. In this particular case, that dataset contains only a few rows describing automobile models and some of their specifications like Horse Power and expected Miles per Gallon.

### Web Interface {#starting-spark-web-interface}

Most of the Spark commands are executed from the R console; however, monitoring and analyzing execution is done through Spark's web interface, see Figure \@ref(fig:starting-spark-web). This interface is a web application provided by Spark which can be accessed by running:

```{r starting-spark-web-code}
spark_web(sc)
```

```{r starting-spark-web-shot, echo=FALSE}
invisible(webshot::webshot(
  "http://localhost:4040/",
  "images/starting-spark-web.png",
  cliprect = c(0, 0, 992, 540),
  zoom = 2
))
```
```{r starting-spark-web, eval=TRUE, fig.width=4, fig.align='center', echo=FALSE, fig.cap='Apache Spark Web Interface'}
render_image("images/starting-spark-web.png", "Apache Spark Web Interface")
```

Printing the `cars` dataset collected a few records to be displayed in the R console. You can see in the Spark web interface that a job was started to collect this information back from Spark. You can also select the **storage** tab to see the "mtcars" dataset cached in-memory in Spark, Figure \@ref(fig:starting-spark-web-storage).

```{r starting-spark-web-storage-shot, echo=FALSE}
invisible(webshot::webshot(
  "http://localhost:4040/storage/",
  "images/starting-spark-web-storage.png",
  cliprect = c(0, 0, 992, 260),
  zoom = 2
))
```
```{r starting-spark-web-storage, eval=TRUE, fig.width=4, fig.align='center', echo=FALSE, fig.cap='Apache Spark Web Interface - Storage Tab'}
render_image("images/starting-spark-web-storage.png", "Apache Spark Web Interface - Storage Tab")
```

Notice that this dataset is fully loaded into memory since the **fraction cached** is 100%, you can know exactly how much memory this dataset is using through the **size in memory** column.

The executors tab, Figure \@ref(fig:starting-spark-executors), provides a view of your cluster resources. For local connections, you will find only one compute instance active with only 2GB of memory allocated to Spark and 384MB available for computation. The [tunning](#tunning) chapter you will learn how request more compute instances, resources and learn how memory is allocated.

```{r starting-spark-executors-web-shot, echo=FALSE}
invisible(webshot::webshot(
  "http://localhost:4040/executors/",
  "images/starting-spark-web-executors.png",
  cliprect = c(0, 0, 992, 700),
  zoom = 2,
  delay = 3
))
```
```{r starting-spark-executors, eval=TRUE, fig.width=4, fig.align='center', echo=FALSE, fig.cap='Apache Spark Web Interface - Executors Tab'}
render_image("images/starting-spark-web-executors.png", "Apache Spark Web Interface - Executors Tab")
```

The last tab to explore is the environment tab, Figure \@ref(fig:starting-spark-environment), this tab lists all the settings for this Spark application which the [tunning](#tunning) will also introduce them in detail. As you will learn, most settings don't need to be configured explicitly, but in order to properly run at scale, you will have to become familiar with some of them, eventually.

```{r starting-spark-environment-shot, echo=FALSE}
invisible(webshot::webshot(
  "http://localhost:4040/environment/",
  "images/starting-spark-web-environment.png",
  cliprect = c(0, 0, 992, 740),
  zoom = 2
))
```
```{r starting-spark-environment, eval=TRUE, fig.width=4, fig.align='center', echo=FALSE, fig.cap='Apache Spark Web Interface - Environment Tab'}
render_image("images/starting-spark-web-environment.png", "Apache Spark Web Interface - Environment Tab")
```

### Analysis {#starting-analysis}

When using Spark from R to analyze data, you can use SQL (Structured Query Language) or `dplyr` (a grammar of data manipulation). SQL can be used through the `DBI` package; for instance, to count how many records are available in our `cars` dataset we can run:

```{r}
library(DBI)
dbGetQuery(sc, "SELECT count(*) FROM mtcars")
```
```
  count(1)
1       32
```

When using `dplyr`, you write less code and it's often much easier to write than SQL; which is why we won't make use SQL in this book; however, if you are profficient in SQL, this is a viable option to you. For instance, counting records in `dplyr` is more compact and easier to understand.

```{r}
library(dplyr)
count(cars)
```
```
# Source: spark<?> [?? x 1]
      n
  <dbl>
1    32
```

In general, we usually start by analysing data in Spark with `dplyr`, followed by sampling rows and selecting a subset of the available columns, the last step is to collect data from Spark to perform further data processing in R, like data visualization. Let's perform a very simple data analysis example by selecting, sampling and plotting the `cars` dataset in Spark:

```{r}
select(cars, hp, mpg) %>%
  sample_n(100) %>%
  collect() %>%
  plot()
```
```{r echo=FALSE}
select(cars, hp, mpg) %>%
  sample_n(100) %>%
  collect() %>%
  ggplot() +
    geom_point(aes(hp, mpg), size = 2.5, color = "#555555") +
    plot_style() +
    labs(title = "Vehicle Efficiency", subtitle = "Miles per gallon vs horse power") +
    geom_hline(yintercept = 0, size = 1, colour = "#333333") +
    scale_x_continuous(name = "horse power",
                       limits = c(50, 360),
                       breaks = seq(50, 400, by = 50),
                       labels = c("50","100", "150", "200", "250", "300", "350", "400")) +
    scale_y_continuous(name = "miles per gallon") +
    ggsave("images/starting-cars-hp-vs-mpg.png", width = 10, height = 5)
```
```{r starting-cars-hp-vs-mpg, eval=TRUE, echo=FALSE, fig.cap='Horse Power vs Miles per Gallon', fig.align = 'center'}
render_image("images/starting-cars-hp-vs-mpg.png", "Horse Power vs Miles per Gallon")
```

The plot in Figure \@ref(fig:starting-cars-hp-vs-mpg), shows that as we increase the horse power in a vehicle, their fuel efficiency measured in miles per gallon gets reduced. While this is insightful, it's hard to predict numerically how increased horse power would affect fuel effiency, modeling can help us overcome this.

### Modeling {#starting-modeling}

While data analysis can take you quite far when understanding data, building a mathematical model that describes and generalizes the dataset is quite powerful. In the [introduction](#intro) chapter you learned that the fields of machine learning and data science make use of mathematical models to perform predictions and find additional insights.

For instance, we can use a linear model to approximate the relationship between fuel efficiency and horse power:

```{r}
model <- ml_linear_regression(cars, mpg ~ hp)
```

This model can now be used to predict values that are not in the original datset. For instance, we can add entries for cars with horse power beyond 250 and also visualize the predicted values as shown in Figure \@ref(fig:starting-cars-hp-vs-mpg-prediction).

```{r}
model %>%
  ml_predict(copy_to(sc, data.frame(hp = 250 + 10 * 1:10))) %>%
  transmute(hp = hp, mpg = prediction) %>%
  full_join(select(cars, hp, mpg)) %>%
  collect() %>%
  plot()
```
```{r echo=FALSE}
model %>%
  ml_predict(copy_to(sc, data.frame(hp = 250 + 10 * 1:10))) %>%
  transmute(hp = hp, mpg = prediction, label = "Predicted") %>%
  full_join(transmute(cars, hp = hp, mpg = mpg, label = "Original")) %>%
  collect() %>%
  ggplot() +
    geom_point(aes(hp, mpg, color = label), size = 2.5) +
    scale_color_grey(start = 0.6, end = 0.2) +
    plot_style() +
    labs(title = "Vehicle Efficiency", subtitle = "Miles per gallon vs horse power with predictions") +
    geom_hline(yintercept = 0, size = 1, colour = "#333333") +
    scale_x_continuous(name = "horse power",
                       limits = c(50, 360),
                       breaks = seq(50, 350, by = 50),
                       labels = c("50","100", "150", "200", "250", "300", "350")) +
    scale_y_continuous(name = "miles per gallon") +
    ggsave("images/starting-cars-hp-vs-mpg-prediction.png", width = 10, height = 5)
```
```{r starting-cars-hp-vs-mpg-prediction, eval=TRUE, echo=FALSE, fig.cap='Horse power vs miles per gallon with predictions', fig.align = 'center'}
render_image("images/starting-cars-hp-vs-mpg-prediction.png", "Horse power vs miles per gallon with predictions")
```

In addition, we can refine our insights using the `broom` package to retrive additional statistics from our model that can help us asses it's quality.

```{r}
broom::glance(model)
```
```
# A tibble: 1 x 5
  explained.varia… mean.absolute.e… mean.squared.er… r.squared
             <dbl>            <dbl>            <dbl>     <dbl>
1             21.2             2.91             14.0     0.602
# … with 1 more variable: root.mean.squared.error <dbl>
```

While the previous example lacks many of the appropriate techniques you should use while modeling, it's also a simple example to briefly introduce the modeling capabilities of Spark. All the Spark models, techniques and best practices will be properly introduced in the [modeling](#modeling) chapter.

### Data {#starting-data}

For simplicity, we copied the `mtcars` dataset into Spark; however, data is usually not copied into Spark. Instead, data is read from existing data sources in a variety of formats like plain text, CSV, JSON, JDBC and many more which, the [data](#data) chapter will introduce in detail. For instance, we can export our `cars` dataset as a CSV file:

```{r}
spark_write_csv(cars, "cars.csv")
```

In practice, we would read an existing dataset from a distributed storage system like HDFS, but we can also read back from the local file system:

```{r}
cars <- spark_read_csv(sc, "cars.csv")
```

### Extensions {#starting-extensions}

In the same way that R is known for it's vibrant community of package authors, at a smaller scale, many extensions for Spark and R have been written and are available to you. The [extensions](#extensions) chapter will introduce many interesting ones to perform advanced modeling, graph analysis, preprocess datasets for deep learning, etc.

For instance, the `sparkly.nested` extension is an R package that extends `sparklyr` to help you manage values that contain nested information. A common use case arises while dealing with JSON files which contain nested lists that require preprocessing before doing meaningful data analysis. To use this extension, we have to first install it as follows:

```{r}
install.packages("sparklyr.nested")
```

Then we can use this extension to group all the horse power data points over the number of cylinders:

```{r}
sparklyr.nested::sdf_nest(cars, hp) %>%
  group_by(cyl) %>%
  summarize(data = collect_list(data))
```
```
# Source: spark<?> [?? x 2]
    cyl data       
  <int> <list>     
1     6 <list [7]> 
2     4 <list [11]>
3     8 <list [14]>
```

While nesting data makes it harder to read, it is a requirement while dealing with nested data formats like JSON using the `spark_read_json()` and `spark_write_json()` functions.

### Distributed R {#starting-distributed-r}

For those few cases where a particular functionality is not available in Spark and no extension has been developed, you can consider distributing your own R code across the Spark cluster. This is a powerful tools but comes with additional complexity that you should only use as a last resort option.

Suppose that we need to round all the values across all the columns in our dataset, one approach would be running custom R code making use of R's `round()` function:

```{r}
cars %>% spark_apply(~round(.x))
```
```
# Source: spark<?> [?? x 11]
     mpg   cyl  disp    hp  drat    wt  qsec    vs    am  gear  carb
   <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
 1    21     6   160   110     4     3    16     0     1     4     4
 2    21     6   160   110     4     3    17     0     1     4     4
 3    23     4   108    93     4     2    19     1     1     4     1
 4    21     6   258   110     3     3    19     1     0     3     1
 5    19     8   360   175     3     3    17     0     0     3     2
 6    18     6   225   105     3     3    20     1     0     3     1
 7    14     8   360   245     3     4    16     0     0     3     4
 8    24     4   147    62     4     3    20     1     0     4     2
 9    23     4   141    95     4     3    23     1     0     4     2
10    19     6   168   123     4     3    18     1     0     4     4
# … with more rows
```

If you are a profficient R user, it can be quite tempting to use `spark_apply()` for everything, but please, don't! `spark_apply()` was designed for advanced use cases where Spark falls short; instead, you will learn how to do proper data analysis and modeling without having to distribute custom R code across your cluster.

### Streaming {#starting-streaming}

While processing large static datasets is the most typical use case for Spark, processing dynamic datasets in realtime is also possible and for some applications, a requirement. You can think of a streaming dataset as a static data source with new data arriving continously, like stock market quotes. Streaming data is usually read from Kafka (an open-source stream-processing software platform) or from distributed storage that receives new data continuously.

To try out streaming, lets first  create an `input/` folder with some data that we will use as the input for this stream:

```{r}
dir.create("input")
write.csv(mtcars, "input/cars_1.csv", row.names = F)
```

Then we will define a stream that processes incoming data from the `input/` folder, performs a custom transformation in R and, pushes the output into an `output/` folder

```{r}
stream_read_csv(sc, "input/") %>%
  spark_apply(~sapply(.x, jitter)) %>%
  stream_write_csv("output/")
```
```
Stream: 720aac2a-d4aa-4e6c-828d-325d8b017fdb
Status: Waiting for next trigger
Active: TRUE
```

As soon as the stream of realtime data starts, the `input/` folder is processed and turned into a set of new files under the `output/` folder containing the new transformed files. Since the input contained only one file, the output folder will also contain a single file resulting from applying the custom `spark_apply()` transformation.

```{r}
dir("output", pattern = ".csv")
```
```
[1] "part-00000-eece04d8-7cfa-4231-b61e-f1aef8edeb97-c000.csv"
```

Up to this point, this resembles static data processing; however, we can keep adding files to the `input/` location and Spark will parallelize and process data automatically. Let's add one more file and validate that it's automatically processed.

```{r}
# Write more data into the stream source
write.csv(mtcars, "input/cars_2.csv", row.names = F)

# Wait for the input stream to be processed
Sys.sleep(1)

# Check the contents of the stream destination
dir("output", pattern = ".csv")
```
```
[1] "part-00000-2d8e5c07-a2eb-449d-a535-8a19c671477d-c000.csv"
[2] "part-00000-eece04d8-7cfa-4231-b61e-f1aef8edeb97-c000.csv"
```

You can use `dplyr`, SQL, Spark models or distributed R to analyze streams in realtime, we will properly introduce you to all the interesting transformations you can perform to analyze realtime data during the [streaming](#streaming) chapter.

### Logs {#starting-logs}

Logging is definetely less interesting that realtime data processing; however, it's a tool you should be familiar with. A log is just a text file where Spark will append information relevant to the execution of tasks in the cluster. For local clusters, we can retrieve all the recent log by running:

```{r starting-logs}
spark_log(sc)
```
```
18/10/09 19:41:46 INFO Executor: Finished task 0.0 in stage 5.0 (TID 5)...
18/10/09 19:41:46 INFO TaskSetManager: Finished task 0.0 in stage 5.0...
18/10/09 19:41:46 INFO TaskSchedulerImpl: Removed TaskSet 5.0, whose...
18/10/09 19:41:46 INFO DAGScheduler: ResultStage 5 (collect at utils...
18/10/09 19:41:46 INFO DAGScheduler: Job 3 finished: collect at utils...
```

Or we can retrieve specific log entries containing, say `sparklyr`, by using the `filter` parameter as follows:

```{r starting-logs-filter}
spark_log(sc, filter = "sparklyr")
```
```
## 18/10/09 18:53:23 INFO SparkContext: Submitted application: sparklyr
## 18/10/09 18:53:23 INFO SparkContext: Added JAR...
## 18/10/09 18:53:27 INFO Executor: Fetching spark://localhost:52930/...
## 18/10/09 18:53:27 INFO Utils: Fetching spark://localhost:52930/...
## 18/10/09 18:53:27 INFO Executor: Adding file:/private/var/folders/...
```

Most of the time, you won't need to worry about Spark logs, except in cases where you need to troubleshoot a failed computation; in those cases, logs are an invaluable resource to be aware of, now you know.

## Disconnecting {#starting-disconnecting}

For local clusters (really, any cluster) once you are done processing data you should disconnect by running:

```{r starting-disconnect}
spark_disconnect(sc)
```

This will terminate the connection to the cluster as well as the cluster tasks . If multiple Spark connections are active, or if the conneciton instance `sc` is no longer available, you can also disconnect all your Spark connections by running:

```{r starting-disconnect-all}
spark_disconnect_all()
```

Notice that exiting R, RStudio or restarting your R session will also cause the Spark connection to terminate, which in turn terminates the Spark cluster and cached data that is not explicitly persisted.

## Using RStudio {#starting-using-spark-from-rstudio}

Since it's very common to use RStudio with R, `sparklyr` provides RStudio extensions to help simplify your workflows and increase your productivity while using Spark in RStudio. If you are not familiar with RStudio, take a quick look at the [Using RStudio](#appendix-using-rstudio) appendix section. Otherwise, there are a couple extensions worth highlighting.

First, instead of starging a new connections using `spark_connect()` from RStudio's R console, you can use the **new connection** action from the **connections pane** and then, select the Spark connection which will open the dialog shown in Figure \@ref(fig:starting-rstudio-new-connection). You can then customize the versions and connect to Spark which will simply generate the right `spark_connect()` command and execute this in the R console for you.

```{r starting-rstudio-new-connection, eval=TRUE, fig.width=4, fig.align='center', echo=FALSE, fig.cap='RStudio New Spark Connection'}
render_image("images/starting-rstudio-new-spark-connection.png", "RStudio New Spark Connection")
```

Second, once connected to Spark, either by using the R console or through RStudio's connections pane, RStudio will display your datasets available in the connections pane, see Figure \@ref(fig:starting-rstudio-connections-pane). This is a useful way to track your existing datasets and provides an easy way to explore each of them.

```{r starting-rstudio-connections-pane, eval=TRUE, fig.width=4, fig.align='center', echo=FALSE, fig.cap='RStudio Connections Pane'}
render_image("images/starting-rstudio-connections-pane.png", "RStudio Connections Pane")
```

Additionally, an active connection provides the following custom actions:

- **Spark**: Opens the Spark web interface, a shortcut to `spark_ui(sc)`.
- **Log**: Opens the Spark web logs, a shortcut to `spark_log(sc)`.
- **SQL**: Opens a new SQL query, see `DBI` and SQL support in the data [analysis](#analysis) chapter.
- **Help**: Opens the reference documentation in a new web browser window.
- **Disconnect**: Disconnects from Spark, a shortcut to `spark_disconnect(sc)`.

The rest of this book will use plain R code, it is up to you to execute this code in the R console, RStudio, Jupyter Notebooks or any other tool that support executing R code since, the code provided in this book executes in any R environment.

## Resources {#starting-resources}

While we've put significant effort into simplifying the onboarding process, there are many additional resources that can help you troubleshoot particular issues while getting started and, in general, introduce you to the broader Spark and R communities to help you get specific answers, discuss topics and get connected with many users actevely using Spark with R. 

- **Documentation**: This should be your first stop to learn more about Spark when using R. The documentation is kept up to date with examples, reference functions and many more relevant resources, [spark.rstudio.com](https://spark.rstudio.com).
- **Blog**: To keep up to date with major `sparklyr` announcements, you can follow the RStudio blog, [blog.rstudio.com/tags/sparklyr](https://blog.rstudio.com/tags/sparklyr).
- **Community**: For general `sparklyr` questions, you can post then in the RStudio Community tagged as `sparklyr`, [community.rstudio.com/tags/sparklyr](https://community.rstudio.com/tags/sparklyr).
- **Stack Overflow**: For general Spark questions, Stack Overflow is a great resource, [stackoverflow.com/questions/tagged/apache-spark](https://stackoverflow.com/questions/tagged/apache-spark); there are also many topics specifically about `sparklyr`, [stackoverflow.com/questions/tagged/sparklyr](https://stackoverflow.com/questions/tagged/sparklyr).
- **Github**: If you believe something needs to be fixed, open a GitHub issue or send us a pull request, [github.com/rstudio/sparklyr](https://github.com/rstudio/sparklyr).
- **Gitter**: For urgent issues, or to keep in touch, you can chat with us in Gitter, [gitter.im/rstudio/sparklyr](https://gitter.im/rstudio/sparklyr).

## Recap {#starting-recap}

In this chapter you learned about the prerequisites required to work with Spark, how to connect to Spark using `spark_connect()`, install a local cluster using `spark_install()`, load a simple dataset, launch the web interface and display logs using `spark_web(sc)` and `spark_log(sc)` respectively, disconnect from RStudio using `spark_disconnect()` and we closed this chapter presenting the RStudio extensions `sparklyr` provides.

It is our hope that this chapter will help anyone interested in learning cluster computing using Spark and R getting started, ready to experiment on your own and ready to tackle actual data analysis and modeling problems which, the next two chapters will introduce you. The next chapter, analysis, will present data analysis as the process to inspect, clean, and transform data with the goal of discovering useful information. Modeling can be considered part of data analysis; however, it deserves it's own chapter to truly understand and take advantage of the modeling functionality available in Spark.
