```{r include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
source("r/render.R")
source("r/plots.R")
library(ggplot2)
```

# Streaming {#streaming}

As the velocity of the generation of data increases, so does the need to real-time analysis. Real-time refers to the ability to continuously analyze data from a constantly updating data feed. This is usually called Stream Analytics. In Stream Analytics, data is analyzed as its generated, not retroactively as it happens with "everyday" analysis.  

This chapter will cover how to analyze a stream of data using R and Spark. It will also cover basics about stream analysis, along with how to implement these using `sparklyr` and other R packages.

## Spark Streaming 

Spark Streaming is an extension of the core Spark API.  It is used for processing live streams of data.  It does this in a scalable, high-throughput, and fault tolerant way. It also allows for the current data to be joined with the historical data.

Spark Streaming works by splitting the live input into small batches.  Each batch is processed by Spark individually.  The output from Spark is also in small batches.  This process is not visible to the user.  Spark displays streams as a DStream (discretize stream).  The DStream represents the small batches as one continuous stream. Inside Spark, the DStream is represented as a sequence of Resilient Distributed Datasets (RDD).

The best resource to learn how Spark analyzes streams is the Apache Spark's Official site [@streaming-programming-guide]. This chapter will cover just enough Spark Streaming concepts to help you understand the mechanics of what the R code is doing.  It is recommended to read the official resource, specially if you need to implement solutions based on Spark Streaming.

## Working with Spark Streams

In practice, a Spark stream update is a three stage operation.  Two of the stages are driven by new files being added, either from the stream, or as a result of the Spark analysis. This is the breakdown of the three stages:

1. **Read** - The stream is expected to append new files in a specified folder. Those new files contain the most recent information from the stream.  Spark monitors the folder, and reads the data from the files. The following formats are supported: CSV, text, JSON, parquet, Kafka, JDBC, and orc.

1. **Transform** - Spark applies the desired operations on top of the data. No special `sparklyr` functions are needed to transform stream data.  You can use same `dplyr` verbs, Spark transformers and even native R code (via `spark_apply()`).

1. **Write** - The results of the transformed input are saved in a different folder.  The following file formats are supported: CSV, text, JSON, parquet, Kafka, JDBC, and orc.

Figure 11.1 provides a visual aid to what each stage does and how they connect:

```{r echo=FALSE, out.width='100%', out.height='220pt', fig.cap='Working with Spark Streams', fig.align = 'center', eval = TRUE}
render_nomnoml("
#direction: right
#padding: 10
#fontSize: 14
#leading: 2
[<note> Input
Folder] -> [<transceiver> Spark reads data 
from files in folder] 
[Spark reads data 
from files in folder] -> [<transceiver> Spark applies 
the operations] 
[Spark applies 
the operations] -> [<transceiver> Spark writes 
results to folder]
 [<transceiver> Spark writes 
results to folder] -> [<note> Output 
Folder] 
[Spark reads data 
from files in folder] -/- [<label> 2. Transform]
[Input
Folder] -/- [<label> 1. Read]
[Spark applies 
the operations] -/- [<label> 3. Write]
", "images/streaming-working.png")
```

## Read/Write functions

Here is the breakdown of the available `sparklyr` functions for reading and writing: 

| Format  | Read                | Write                |
|---------|---------------------|----------------------|
| CSV     | stream_read_csv     | stream_write_csv     |
| JSON    | stream_read_json    | stream_write_json    |
| Kafka   | stream_read_kafka   | stream_write_kafka   |
| ORC     | stream_read_orc     | stream_write_orc     |
| Parquet | stream_read_parquet | stream_write_parquet |
| Text    | stream_read_text    | stream_write_text    |
| Memory  |                     | stream_write_memory  |

In the same way all of the read and write operations in `sparklyr` for Spark Standalone, or in `sparklyr`’s local mode, the input and output folders are actual Operating System file system folders. For YARN managed clusters, these will be folder locations inside the Hadoop File System (HDFS). 

## Additional features

The `sparklyr` package goes beyond providing an easy-to-use-interface to work with all three stages of Spark streaming.  The package includes additional features which provide a more complete integration with R:

1. An out-of-the box graph visualization to monitor the stream.

1. Stream generator for testing and learning purposes.

1. A Shiny reactive function. It allows Shiny apps to read the contents of a steam.

### Stream monitor

The `stream_view()` function will generate a Shiny app which displays the current state, as well as the history, of the stream.  An example of how to use it is available in the **Intro Example** section.

### Stream generator

The `stream_generate_test()` function creates a local test stream.  This function works independently from a Spark connection. The following example will create five files in sub-folder called "source".  The files will be created one second apart from the previous file's creation. 

```{r}
library(sparklyr)

stream_generate_test(iterations = 5, path = "source", interval = 1)
```

After the function completes, all of the files should show up in the "source" folder. Notice that the file size vary.  This is so that it simulates what a true stream would do.  

```{r}
file.info(file.path("source", list.files("source")))[1] 
```
```
##                     size
## source/stream_1.csv   44
## source/stream_2.csv  121
## source/stream_3.csv  540
## source/stream_4.csv 2370
## source/stream_5.csv 7236
```
The `stream_generate_test()` by default will create a single numeric variable data frame.  

```{r}
readr::read_csv("source/stream_5.csv")
```
```
## # A tibble: 1,489 x 1
##        x
##    <dbl>
##  1   630
##  2   631
##  3   632
##  4   633
##  5   634
##  6   635
##  7   636
##  8   637
##  9   638
## 10   639
## # ... with 1,479 more rows
```

### Shiny reactive

The "live" nature of the Spark stream goes along well with the reactivity of Shiny.  The idea is that your Shiny app can automatically display the latest results as fast as Spark can process them.  The `reactiveSpark()` function provides that integration.  

## Intro example

This section will use a very simple example to introduce the mechanics of Spark Streaming, and how `sparklyr` interacts with it. This is very simple example. It will only move the input contents to the output contents without any transformations being done to it. 

1. Open a local Spark session
    ```{r}
    sc <- spark_connect(master = "local")
    ```

1. Remove the "source" and "destination" folders.  This step ensures a clean slate if you try to run the example again.
    ```{r}
    if(file.exists("source")) unlink("source", TRUE)
    if(file.exists("destination")) unlink("destination", TRUE)
    ```

1. Just like with `read_csv()`, `stream_read_csv()` needs a file specification. To save ourselves from providing one, a single test file is generated.
    ```{r}
    stream_generate_test(iterations = 1)
    ```

1. `stream_read_csv()` starts the ingestion part of the job.  It corresponds to the **1. Read** stage described in the previous section.
    ```{r}
    read_folder <- stream_read_csv(sc, "source")
    ```

1. Set the output of the job to read the incoming data. That is done by passing the *read_folder* variable, set in the previous step.  It corresponds to the **3. Read** stage described in the previous section.
    ```{r}
    write_output <- stream_write_csv(read_folder, "destination")
    ```

1. The library`future` will allow the test generation to run in a asynchronous fashion. This is needed because the next step, `stream_view()` will start a Shiny app which takes over the R session. 
    ```{r}
    library(future)
    invisible(future(stream_generate_test(interval = 0.3)))
    ```

1. `stream_view()` is the out-of-the box graph visualization to monitor the stream that was mentioned in the *sparklyr Interface* section.
    ```{r}
    stream_view(write_output)
    ```

The Shiny app shows up in the Viewer pane.  The column bars will slowly accumulate in the app's plot After the test generator completes, the plot should look like what Figure 11.2 shows.

```{r, eval = TRUE, fig.width=10, fig.height=5, fig.cap='stream_view() output', fig.align = 'center', echo = FALSE}
render_image("images/streaming-stream-view.png")
```

The final step is to clean up the stream and Spark connection

```{r}
stream_stop(write_output)
spark_disconnect(sc)
```


## Transformations 

Streams can be transformed using `dplyr`, SQL queries, ML Pipelines or R code. We can use as many transformations as needed in the same way that Spark data frames can be transformed with `sparklyr`. The transformation source can be streams or data frames but the output is always a stream. If needed, one can always take a snapshot from the destination stream and save the output as a data frame, which is what `sparklyr` will do for you if a destination stream is not specified. 

### dplyr

The same `dplyr` verbs can be used on top of a Spark Stream.  The following example shows how easy it is to filter rows and add columns to data from an input folder.

```{r}
sc <- spark_connect(master = "local")
if(file.exists("source")) unlink("source", TRUE)

stream_generate_test(iterations = 5)

stream_read_csv(sc, "source") %>%
  filter(x > 700) %>%
  mutate(y = round(x / 100))
```
```
## # Source: spark<?> [inf x 2]
##        x     y
##    <int> <dbl>
##  1   701     7
##  2   702     7
##  3   703     7
##  4   704     7
##  5   705     7
##  6   706     7
##  7   707     7
##  8   708     7
##  9   709     7
## 10   710     7
## # ... with more rows
```

It also is possible to perform aggregations over the entire history of the stream.  The history could be filtered or not.
```{r}
stream_read_csv(sc, "source") %>%
  filter(x > 700) %>%
  mutate(y = round(x / 100)) %>%
  count(y) 
```
```
## # Source: spark<?> [inf x 2]
##       y     n
##   <dbl> <dbl>
## 1     8   200
## 2     9   200
## 3    10   102
## 4     7    98
```

Grouped aggregations of the latest data in the stream require a time stamp.  The time stamp will be of when reading function, in this case `stream_read_csv()` , first "saw" that specific record. In Spark stream terms, that time stamp is called a "watermark".  The `spark_watermark()` function is used to add the time stamp.  For this exercise, the watermark will be the same for all records.  That is because the 5 files were read by the stream after they were created.  Please note that only Kafka and memory *outputs* support watermarks.  

```{r}
stream_read_csv(sc, "source") %>%
  stream_watermark()
```

```
## # Source: spark<?> [inf x 2]
##        x timestamp          
##    <int> <dttm>             
##  1   630 2019-04-07 15:44:50
##  2   631 2019-04-07 15:44:50
##  3   632 2019-04-07 15:44:50
##  4   633 2019-04-07 15:44:50
##  5   634 2019-04-07 15:44:50
##  6   635 2019-04-07 15:44:50
##  7   636 2019-04-07 15:44:50
##  8   637 2019-04-07 15:44:50
##  9   638 2019-04-07 15:44:50
## 10   639 2019-04-07 15:44:50
## # ... with more rows
```

After the watermark is created, it can be used in the `group_by()` verb.  It can then be piped into a `summarise()` function to get some stats of the stream.

```{r}
stream_read_csv(sc, "source") %>%
  stream_watermark() %>%
  group_by(timestamp) %>%
  summarise(
    max_x = max(x, na.rm = TRUE),
    min_x = min(x, na.rm = TRUE),
    count = n()
  ) 
```
```
## # Source: spark<?> [inf x 4]
##   timestamp           max_x min_x count
##   <dttm>              <int> <int> <dbl>
## 1 2019-04-07 15:45:59  1000     1  2122
```

### Transformer functions

Transformer functions can also be used to modify a stream.  They can also be combined with the regular `dplyr` functions.  

```{r}
stream_read_csv(sc, "source") %>%
  mutate(x = as.numeric(x)) %>%
  ft_bucketizer("x", "buckets", splits = 0:10 * 100) %>%
  count(buckets)  %>%
  arrange(buckets)
```

```
## # Source:     spark<?> [inf x 2]
## # Ordered by: buckets
##    buckets     n
##      <dbl> <dbl>
##  1       0   299
##  2       1   220
##  3       2   200
##  4       3   200
##  5       4   200
##  6       5   200
##  7       6   201
##  8       7   200
##  9       8   200
## 10       9   202
```

### R code

Arbitrary R code can also be used to transform a stream with the use of `spark_apply()`. Following the same principles from executing R code over Spark data frames, for structured streams, `spark_apply()` runs R code over each executor in the cluster where data is available, this enables processing high-throughput streams and fulfill low-latency requirements.

```{r}
stream_read_csv(sc, "source") %>%
  spark_apply(~ nrow(.x), list(n="integer"))
```
```
## # Source: spark<?> [inf x 1]
##       n
##   <int>
## 1  1962
## 2   148
## 3    12
```

### ML Pipelines

Spark pipelines can be used for scoring streams, but not to train over streaming data. The former is fully supported while the latter is a feature under active development by the Spark community.

1. In order to try  scoring data in a stream, it is necessary to first create a Pipeline Model.  The following build, fits and saves a simple pipeline. It also opens and closes the Spark connection.
    ```{r}
    sc <- spark_connect(master = "local")
    cars <- copy_to(sc, mtcars, "mtcars_remote")
    sc %>%
      ml_pipeline() %>%
      ft_binarizer("mpg", "over_30",30) %>%
      ft_r_formula(over_30 ~ wt) %>%
      ml_logistic_regression() %>%
      ml_fit(cars) %>%
      ml_save("cars_model")
    spark_disconnect(sc)
    ```

1. A new connection of Spark is opened.  The saved model is loaded into the new connection.  
    ```{r}
    sc <- spark_connect(master = "local")
    model <- ml_load(sc, "cars_model")
    ```

1. Data that can be used for predictions is needed.  The `stream_generate_test()` can be used for this as well.  Instead of relying on the default output, the *mtcars* variable is passed to it.  
    ```{r}
    if(file.exists("source")) unlink("source", TRUE)
    stream_generate_test(mtcars, iterations = 5)
    ```

1. The `ml_transform()` function can now be used on top of the stream. Because the function expects the model as the first function, the piping works a little different. Instead of starting with reading the stream, we start with the model, and use the stream input as the argument on `ml_transform()`
    ```{r}
    model %>%
      ml_transform(stream_read_csv(sc, "source"))
    ```

    ```
    ## # Source: spark<?> [inf x 17]
    ##      mpg   cyl  disp    hp  drat    wt  qsec    vs    am
    ##    <dbl> <int> <dbl> <int> <dbl> <dbl> <dbl> <int> <int>
    ##  1  15.5     8 318     150  2.76  3.52  16.9     0     0
    ##  2  15.2     8 304     150  3.15  3.44  17.3     0     0
    ##  3  13.3     8 350     245  3.73  3.84  15.4     0     0
    ##  4  19.2     8 400     175  3.08  3.84  17.0     0     0
    ##  5  27.3     4  79      66  4.08  1.94  18.9     1     1
    ##  6  26       4 120.     91  4.43  2.14  16.7     0     1
    ##  7  30.4     4  95.1   113  3.77  1.51  16.9     1     1
    ##  8  15.8     8 351     264  4.22  3.17  14.5     0     1
    ##  9  19.7     6 145     175  3.62  2.77  15.5     0     1
    ## 10  15       8 301     335  3.54  3.57  14.6     0     1
    ## # ... with more rows, and 8 more variables: gear <int>,
    ## #   carb <int>, over_30 <dbl>, features <list>,
    ## #   label <dbl>, rawPrediction <list>,
    ## #   probability <list>, prediction <dbl>
    ```

### Shiny integration

The `reactiveSpark()` provides a mechanism to process the transformations on a stream. It allows you to circumvent the need for writing an output.  Also, because it does not depend on the stream writing functions, it is possible to to use watermark groups. 

This section's example will result in a Shiny app. It will start to accumulate and display the current and historical results. The app's output is shown on figure 11.3.

1. Start by opening a Spark connection and begin a test generation.
    ```{r}
    sc <- spark_connect(master = "local")
    if(file.exists("source")) unlink("source", TRUE)
    invisible(future(stream_generate_test(interval = 0.2, iterations = 10)))
    ```

1. Load the `shiny` library and create a simple *UI* function with one table output.
    ```{r}
    library(shiny)
    ui <- function() tableOutput("table")
    ```

1. The *server* function contains a `reactiveSpark()` function.  This function reads the stream, adds the watermark and then performs the aggregation.  The results are then rendered via the *table* output.
    ```{r}
    server <- function(input, output, session){
      ps <- stream_read_csv(sc, "source")  %>%
        stream_watermark() %>%
        group_by(timestamp) %>%
        summarise(
          max_x = max(x, na.rm = TRUE),
          min_x = min(x, na.rm = TRUE),
          count = n()) %>%
        reactiveSpark()  # Spark stream reactive
      output$table <- renderTable(
        ps() %>%
          mutate(timestamp = as.character(timestamp))
      )}
    ```

1. The Shiny app can be activated with `runGadget()`.  
    ```{r}
    runGadget(ui, server)
    ```

```{r, eval = TRUE,  fig.align = 'center', fig.cap = 'Shiny reactive', echo = FALSE}
render_image("images/streaming-shiny-1.png")
```

## Full example

```{r}
if(file.exists("source")) unlink("source", TRUE)
row_numbers <- sample(seq_len(nrow(mtcars)), 1000, replace = TRUE)
mtcars_stream <- mtcars[row_numbers, ]
invisible(future(stream_generate_test(mtcars_stream, interval = 0.2, iterations = 10)))
```


```{r}
sc <- spark_connect(master = "local", config = conf)
model <- ml_load(sc, "cars_model")
read_folder <- stream_read_csv(sc, "source")
process <- model %>%
  ml_transform(read_folder) %>%
  stream_watermark() %>%
  count(timestamp, over_30, prediction)
write_folder <- stream_write_memory(process, "mtcars_predictions")
```


```{r}
library(shiny)
ui <- function() tableOutput("table")
server <- function(input, output, session){
  autoInvalidate <- reactiveTimer(2000)
  output$table <- renderTable({
    autoInvalidate()
    tbl(sc, "mtcars_predictions") %>%
      group_by(over_30, prediction) %>%
      summarise(results = sum(n, na.rm = TRUE)) %>%
      select(actual = over_30, prediction, results)
  })}
runGadget(ui, server)
```

```{r, eval = TRUE,  fig.align = 'center', fig.cap = 'Shiny reactive', echo = FALSE}
render_image("images/streaming-shiny-1.png")
```

