```{r include=FALSE, eval=TRUE}
knitr::opts_chunk$set(eval = FALSE)
library(sparklyr)
library(dplyr)
source("r/render.R")
```

# Tuning {#tuning}

In previous chapters we've assumed that computation within a Spark cluster works efficiently. While this is true in some cases, it is often necessary to have some knowledge of the operations Spark runs internally to fine tune configuration settings that will make computations run efficiently. This chapter will explain how Spark computes data over large datasets and provide details on how to fine-tune its operations.

For instance, you will learn how to request more compute nodes and increase the amount of memory which, if you remember from the [Getting Started](#starting) chapter, defaults to only 2GB in local instances. You will learn how Spark unifies computation through partioning, shuffling and caching. As mentioned a few chapters back, this is the last chapter describing the internals of Spark; once you complete this chapter, we believe that you will have the intermediate Spark skills necessary to be productive at using Spark.

In subsequent chapters, Extensions, Distributed R and Streaming, you will learn exciting techniques to deal with specific modeling, scaling and computation problems. However, we must first understand how spark performs internal computations, what pieces we can control, and why.

## Overview

Spark performs distributed computation by configuring, partitioning, executing, shuffling, caching and serializing data, tasks and resources across multiple machines:

- [**Configuring**](#tuning-configuring) requests the cluster manager for resources: total machines, memory, etc.
- [**Partitioning**](#tuning-configuring) splits the data among various machines. Partitions can be either implicit or explicit.
- [**Executing**](#tuning-configuring) means running an arbitrary transformation over each partition.
- [**Shuffling**](#tuning-configuring) redistributes data to the correct machine.
- [**Caching**](#tuning-configuring) preserves data in-memory across different computation cycles.
- [**Serializing**](#tuning-serializing) transforms data to be sent over the network to other workers or back to the driver node.

To illustrate each concept, we will create three partitions with unordered integers and then sort them with `arrange()`:

```{r echo=FALSE}
library(sparklyr)
library(dplyr)

sc <- spark_connect(master = "local")
```
```{r}
data <- copy_to(sc, 
  data.frame(id = c(4, 9, 1, 8, 2, 3, 5, 7, 6)),
  repartition = 3)

data %>% arrange(id) %>% collect()
```

The diagram in Figure \@ref(fig:tuning-overview) shows how this sorting **job** would conceptually work across a cluster of machines. First, Spark would **configure** the cluster to use three worker machines. In this example, the numbers 1-9 are partitioned across three storage instances. Since the **data** is already partitioned, each worker node loads this implicit **partition**; for instance, `4,9,1` is loaded in the first worker node. Afterwards, a *task* is distributed to each worker to apply a transformation to each data partition in each worker node, this task is denoted by `f(x)`. In this example, `f(x)` **executes** a sorting operation within a partition. Since Spark is general, execution over a partition can be as simple or complex as needed.

The result is then **shuffled** to the correct machine to finish the sorting operation across the entire dataset, this completes a stage. A **stage** is a set of operations that Spark can execute without shuffling data between machines. Once the data is sorted across the cluster, the sorted results can be optionally **cached** in memory to avoid rerunning this computation multiple times. 

Finally, a small subset of the results is **serialized**, through the network connecting the cluster machines, back to the driver node to print a preview of this sorting example.

```{r tuning-overview-code, echo=FALSE}
r2d3::r2d3(
  c(),
  "images/tuning-spark-overview.js",
  dependencies = "images/tuning-diagram.js",
  css = "images/tuning-spark-overview.css"
)
```
```{r tuning-overview, eval=TRUE, echo=FALSE, fig.cap='Sorting Distributed Data with Apache Spark', fig.align='center'}
render_image("images/tuning-spark-overview.png", "Sorting Distributed Data with Apache Spark")
```

Notice that while Figure \@ref(fig:tuning-overview) describes a sorting operation, a similar approach applies to filtering or joining datasets and analyzing and modeling data at scale. Spark provides support to perform custom partitions, custom shuffling, etc; however, most of these lower level operations are not exposed in `sparklyr`; instead, `sparklyr` makes those operations available through higher level commands provided by data [analysis](#analysis) tools like `dplyr` or `DBI`, [modeling](#modeling), and by using many [extensions](#extensions). For those few cases where you might need to implement lowe-level operations, you can always use the Spark's Scala API through an `sparklyr` [extensions](#extension) or run custom [distributed R](#distributed) code.

In order to effectively tune Spark, we will start by getting familiar with Spark's computation [**graph**](#tuning-graph-visualization) and Spark's event [**timeline**](#tuning-event-timeline). Both are accessible through [Spark's web interface](#starting-spark-web-interface).

### Graph {#tuning-graph-visualization}

Spark describes all computation steps using a Directed Acyclic Graph (DAG), which means that all computations in Spark move computation forward without repeating previous steps, this helps Spark optimize computations effectively.

The best way to understand Spark's computation graph for a given operation, sorting for our example, is to open the last **completed query** under the **SQL** tab in [Spark's web interface](#starting-spark-web-interface). Figure \@ref(fig:tuning-graph-sql-render) show the resulting graph for this sorting operation, this graph contains the following operations:

- **WholeStageCodegen:** This block describes that the operations it contains were used to generate computer code that was efficiently translated to **byte code**. There is usually a small cost associated with translating operations into byte code, but this is a small cost to pay since then the operations can be execute much faster from Spark. In general, you can ignore this blog and focus on the blocks that it contains.
- **InMemoryTableScan:** This means that the original dataset `data` was stored in-memory and traversed row-by-row once.
- **Exchange:** Partitions were exchanged, read shuffled, across executors in your cluster.
- **Sort:** Once the records arrived at the right executor, they were sorted in this final stage.

```{r tuning-graph-sql, echo=FALSE}
webshot::webshot(
  "http://localhost:4040/SQL/execution/?id=7",
  file = "images/tuning-spark-graph-visualization-sql.png",
  selector = c("#plan-viz-graph"))

```
```{r tuning-graph-sql-render, eval=TRUE, echo=FALSE, fig.align = 'center', fig.cap='Spark Graph for a Sorting Query'}
render_image("images/tuning-spark-graph-visualization-sql.png", "Spark Graph for a Sorting Query")
```

From the query details, you can then open the last Spark job to arrive to the job details page, you can then expand "DAG Visualization" to visualize a graph similar to Figure \@ref(fig:tuning-graph-render). This graph shows a few additional details and the stages in this job. Notice that there are no arrows pointing back to previous steps, since Spark makes use of acyclic graphs.

```{r tuning-graph, echo=FALSE}
webshot::webshot(
  "http://localhost:4040/jobs/job/?id=4",
  file = "images/tuning-spark-graph-visualization.png",
  selector = c("#plan-viz-graph"))

```
```{r tuning-graph-render, eval=TRUE, echo=FALSE, fig.align = 'center', fig.cap='Spark Graph for a Sorting Job'}
render_image("images/tuning-spark-graph-visualization.png", "Spark Graph for a Sorting Job")
```

Next, we will dive into a Spark stage and explore its event timeline.

### Timeline {#tuning-event-timeline}

The **event timeline** is a great summary of how Spark is spending computation cycles over each stage. Ideally, you want to see this timeline consisting of mostly CPU usage since other tasks can be considered overhead. You also want to see Spark using all the CPUs across all the cluster nodes available to you.

Select the first stage in the current job and expand the "Event Timeline", you should see a timeline similar to Figure \@ref(fig:tuning-timeline-simple). Notice that we explicitly requested three partitions, which is represented by three lanes in this visualization.

```{r tuning-timeline-simple-webshot, echo=FALSE}
webshot::webshot(
  "http://localhost:4040/stages/stage/?id=6&attempt=0",
  file = "images/tuning-spark-event-timeline.png",
  eval = "
  casper.waitForSelector(
    '#task-assignment-timeline',
    function() {
      this.click('.expand-task-assignment-timeline');
    }
  );",
  selector = c(".legend-area", ".table-bordered")
  )
```
```{r tuning-timeline-simple, eval=TRUE, echo=FALSE, fig.align='center', fig.cap='Spark Event Timeline'}
render_image("images/tuning-spark-event-timeline.png", "Spark event timeline")
```

Since our machine is equiped with four CPUs, we can parallelize this computation even further by explicitly repartition data using `sdf_repartition()`:

```{r tuning-timeline-sort-repartition}
data %>% sdf_repartition(4) %>% arrange(id) %>% collect()
```

```{r tuning-timeline-webshot, echo=FALSE}
webshot::webshot(
  "http://localhost:4040/stages/stage/?id=8&attempt=0",
  file = "images/tuning-spark-event-timeline-repartition.png",
  eval = "
  casper.waitForSelector(
    '#task-assignment-timeline',
    function() {
      this.click('.expand-task-assignment-timeline');
    }
  );",
  selector = c(".legend-area", "#task-summary-table"))
```
```{r tuning-timeline-repartition, eval=TRUE, echo=FALSE, fig.align='center', fig.cap='Spark Event Timeline with Additional Partitions'}
render_image("images/tuning-spark-event-timeline-repartition.png", "Spark Event Timeline with Additional Partitions")
```

Figure \@ref(fig:tuning-timeline-repartition) now shows four execution lanes with most time spent under "Executor Computing Time", which shows us that this particular operation is making better use of our compute resources. When working with clusters, requesting more compute nodes from your cluster should shorten computation time. In contrast, for timelines that show significant time spent shuffling, requesting more compute nodes might not shorten time and might actualy makes everything slower. There is no concrete set of rules to follow to optimize a stage; however, as you gain experience understanding this timeline over multiple operations -- you will develop insights as to how to properly optimize Spark operations.

## Configuring {#tuning-configuring}

When tuning a Spark application, the most common resources to configure are memory and cores, specifically:

- **Memory in Driver:** The amount of memory required in the driver node.
- **Memory per Worker:** The amount of memory required in the worker nodes.
- **Cores per Worker:** The number of CPUs to required in the worker nodes.
- **Number of Workers:** The number of workers required for this session.

**Note:** It is recommended to request significantly more memory for the driver than the memory available over each worker node. In most cases, you will want to request one core per worker.

In local mode there are no workers, but we can still configure memory and cores to use through:

```{r tuning-config-local}
# Initialize configuration with defaults
config <- spark_config()

# Memory
config["sparklyr.shell.driver-memory"] <- "2g"

# Cores
config["sparklyr.connect.cores.local"] <- 2

# Connect to local cluster with custom configuration
sc <- spark_connect(master = "local", config = config)
```

When using the Spark Standalone and the Mesos cluster managers, all the available memory and cores are assigned by default; therefore, there are no additional configuration changes required, unless, you want to restrict resources to allow multiple users to share this cluster. In this case you can use `total-executor-cores` to restrict the total executors requested. The "Spark Standalone"[@tuning-spark-standalone] and "Spark on Mesos"[@tuning-spark-on-mesos] guides provided additional information when sharing clusters.

When running under YARN Client, you would configure memory and cores as follows:


```{r}
# Memory in Driver
config["sparklyr.shell.driver-memory"] <- "2g"

# Memory per Worker
config["sparklyr.shell.executor-memory"] <- "2g"

# Cores per Worker
config["sparklyr.shell.executor-cores"] <- 1

# Number of Workers
config["sparklyr.shell.num-executors"] <- 3
```

When using YARN in Cluster mode, `sparklyr.shell.driver-cores` can be used to configure total cores requested in the driver node. The "Spark on YARN"[@tuning-spark-on-yarn] guide provides additional configuration settings worth familiarizing yourself with.

There are a few types of configuration settings:

- **Connect** settings are set as parameters to `spark_connect()`, they are common settings used while connecting.
- **Submit** settings are set while `sparklyr` is being submitted to Spark through `spark-submit`, some dependent on the cluster manager being used.
- **Runtime** settings configure Spark when the Spark session is created, these settings are independent to the cluster manager and specific to Spark.
- **sparklyr** settings configure `sparklyr` behaviour, these settings are independent to the cluster manager and particular to R.

The following subsections present extensive lists of all the available settings. It is not required to fully understand them all while tuning Spark, but skimming through them could prove useful in the future while troubleshooting issues. You can also consider skipping the following settings subsections and use them instead as reference material as needed.

### Connect Settings {#connect-settings}

The following parameters can be used with `spark_connect()`, they configure high-level settings that define the connection method, Spark's intallation path and the version of Spark to use.

| name            | value                                                       |
|-----------------|-------------------------------------------------------------|
| master          | Spark cluster url to connect to. Use "local" to connect to a local instance of Spark installed via `spark_install()`. |
| spark_home      | The path to a Spark installation. Defaults to the path provided by the SPARK_HOME environment variable. If SPARK_HOME is defined, it will be always be used unless the version parameter is specified to force the use of a locally installed version. |
| method          | The method used to connect to Spark. Default connection method is "shell" to connect using spark-submit, use "livy" to perform remote connections using HTTP, or "databricks" when using a Databricks clusters.                |
| app_name        | The application name to be used while running in the Spark cluster.                    |
| version         | The version of Spark to use. Only applicable to "local" Spark connections.             |
| config          | Custom configuration for the generated Spark connection. See spark_config for details. |

You can configure additional settings by specifying a list in the `config` parameter, we will now learn what those settings can be.

### Submit Settings {#submit-settings}

Some settings must be specified when `spark-submit` (the terminal application that launches Spark) is run. For instance, since `spark-submit` launches driver node which runs as a Java instance, choosing how much memory is allocated needs to be specified as a parameter to `spark-submit`.

You can list all the available `spark-submit` parameters by running:

```{r tuning-submit-code}
spark_home_dir() %>% file.path("bin", "spark-submit") %>% system2()
```

For readability, we've provided the output of this command in table format, replacing the `spark-submit` parameter with the appropriate `spark_config()` setting and removing the parameters that are not applicable or already presented in this chapter:

| name                               | value                                                       |
|------------------------------------|-------------------------------------------------------------|
| sparklyr.shell.jars                | Specified as 'jars' parameter in 'spark_connect()'.         |
| sparklyr.shell.packages            | Comma-separated list of maven coordinates of jars to include on the driver and executor classpaths. Will search the local maven repo, then maven central and any additional remote repositories given by 'sparklyr.shell.repositories'. The format for the coordinates should be groupId:artifactId:version. |
| sparklyr.shell.exclude-packages    | Comma-separated list of groupId:artifactId, to exclude while resolving the dependencies provided in 'sparklyr.shell.packages' to avoid dependency conflicts. |
| sparklyr.shell.repositories        | Comma-separated list of additional remote repositories to search for the maven coordinates given with 'sparklyr.shell.packages' |
| sparklyr.shell.files               | Comma-separated list of files to be placed in the working directory of each executor. File paths of these files in executors can be accessed via SparkFiles.get(fileName). |
| sparklyr.shell.conf                | Arbitrary Spark configuration property set as PROP=VALUE.   |
| sparklyr.shell.properties-file     | Path to a file from which to load extra properties. If not specified, this will look for conf/spark-defaults.conf. |
| sparklyr.shell.driver-java-options | Extra Java options to pass to the driver.                   |
| sparklyr.shell.driver-library-path | Extra library path entries to pass to the driver.           |
| sparklyr.shell.driver-class-path   | Extra class path entries to pass to the driver. Note that jars added with 'sparklyr.shell.jars' are automatically included in the classpath. |
| sparklyr.shell.proxy-user          | User to impersonate when submitting the application. This argument does not work with 'sparklyr.shell.principal' / 'sparklyr.shell.keytab'. |
| sparklyr.shell.verbose             | Print additional debug output.                              |

The remaining settings are specific to YARN:

| name                                | value                                                                 |
|-------------------------------------|-----------------------------------------------------------------------|
| sparklyr.shell.queue                | The YARN queue to submit to (Default: "default").                     |
| sparklyr.shell.archives             | Comma separated list of archives to be extracted into the working directory of each executor. |
| sparklyr.shell.principal            | Principal to be used to login to KDC, while running on secure HDFS.   |
| sparklyr.shell.keytab               | The full path to the file that contains the keytab for the principal specified above. This keytab will be copied to the node running the Application Master via the Secure Distributed Cache, for renewing the login tickets and the delegation tokens periodically. |

In general, any `spark-submit` setting is configured through `sparklyr.shell.X`, where `X` is the name of the `spark-submit` parameter without the `--` prefix.

### Runtime Settings

As mentioned, some **Spark** settings configure the session runtime. The runtime settings are a superset of the [submit settings](#submit-settings) since is usually helpful to retrieve the current configuration even if a setting can't be changed.

To list the Spark settings set in your current Spark session, you can run:

```{r tuning-runtime-config}
spark_session_config(sc)
```
```{r tuning-runtime-config-table-code, echo=FALSE}
sc <- spark_connect(master = "local")

settings <- spark_session_config(sc)
cleaned <- gsub("Frameworks/R.framework/Versions/3.5/Resources/library", "...", settings)
cleaned <- gsub("javierluraschi", "...", cleaned)
cleaned <- gsub("^file:", "", cleaned)
saveRDS(data.frame(name = names(settings), value = unlist(unname(cleaned))), "data/tuning-runtime-config.rds")
```
```{r tuning-runtime-config-table, eval=TRUE, echo=FALSE}
knitr::kable(readRDS("data/tuning-runtime-config.rds"))
```

However, there are many more configuration settings available in Spark as described in the "Spark Configuration"[@tuning-spark-configuration] guide. It is not in the scope of this book to describe them all so, if possible, take some time to identify the ones that might be of interest to your paritcular use cases.

### sparklyr Settings

Apart from Spark settings, there are a few settings particular to sparklyr listed below. `sparklyr.connect.cores` is useful to set the CPU cores to use in local mode; the remaining ones are not used as much while tuning, but they can prove helpful while troubleshooting other issues.

```{r tuning-settings-sparklyr, eval=FALSE}
spark_config_settings()
```
```{r tuning-settings-sparklyr-code, eval=TRUE, echo=FALSE}
knitr::kable(spark_config_settings())
```

## Partitioning {#tuning-partitioning}

As mentioned in the [introduction](#intro-background) chapter, MapReduce and Spark were designed with the purpose of performing computations against data stored across many machines. The subset of the data available for computation over each compute instance is known as a **partition**.

By default, Spark will compute over each existing **implicit** partition since it's more effective to run computations were the data is already located. However, there are cases where you will want to set an **explicit** partition to help Spark use more efficient use of your cluster resources.

### Implicit

You learned in the [Data](#data) chapter that Spark can read data stored in many formats and different storage systems; however, since shuffling data is an expensive operations, Spark executes tasks reusing the partitions in the storage system. Therefore, these partitions are implicit to Spark since they are already well defined and expensive to rearrange.

There is always an implicit partition for every computation in Spark defined by the distributed storage system, by creating a dataset or by copying datasets into Spark.

You can explore the number of partitions a computation will require through `sdf_num_partitions()`:

```{r tuning-partitioning-implicit}
sdf_len(sc, 10) %>% sdf_num_partitions()
```
```
[1] 2
```

While in most cases the default partitions works just fine, there are cases where we you will need to be explicit on the partitions you choose.

### Explicit

There will be times when you have many more compute instances than data partitions, or much less compute instances than the number of partitions in your data. In both cases, it can help to **repartition** data to match your cluster resources.

Various [data](#data) functions, like `spark_read_csv()`, already support a `repartition` parameter to request Spark to repartition data appropriately. For instance, we can create a sequence of 10 numbers partitioned by 10 as follows:

```{r tuning-partitioning-explicit}
sdf_len(sc, 10, repartition = 10) %>% sdf_num_partitions()
```
```
[1] 10
```

For datasets that are already partitioned, we can also use `sdf_repartition`:

```{r tuning-partitioning-explicit-repartition}
sdf_len(sc, 10, repartition = 10) %>% 
  sdf_repartition(4) %>%
  sdf_num_partitions()
```
```
[1] 4
```

The number of partitions usually significantly changes the speed and resources being used; for instance, the following example calculates the mean over 10M rows with different partition sizes.

```{r tuning-partitioning-explicit-code-run, echo=FALSE}
library(microbenchmark)
library(ggplot2)

partitions_plot <- microbenchmark(
    "1 Partition(s)" = sdf_len(sc, 10^7, repartition = 1) %>%
      summarise(mean(id)) %>% collect(),
    "2 Partition(s)" = sdf_len(sc, 10^7, repartition = 2) %>%
      summarise(mean(id)) %>% collect(),
    times = 10
) %>% 
  autoplot() + theme_light()
ggsave("images/tuning-partition-explicit.png", partitions_plot, width = 10, height = 6)
```
```{r tuning-partitioning-explicit-code}
library(microbenchmark)
library(ggplot2)

microbenchmark(
    "1 Partition(s)" = sdf_len(sc, 10^7, repartition = 1) %>%
      summarise(mean(id)) %>% collect(),
    "2 Partition(s)" = sdf_len(sc, 10^7, repartition = 2) %>%
      summarise(mean(id)) %>% collect(),
    times = 10
) %>% autoplot() + theme_light() 
```
```{r tuning-partitioning-explicit-results, eval=TRUE, echo=FALSE, fig.cap='Computation speed when using explicit Spark partitions', fig.align = 'center'}
render_image("images/tuning-partition-explicit.png", "Computation Speed with Additional Explicit Partitions")
```

The results show that sorting data with two partitions is almost twice as fast; this is the case since two CPUs can be used to execute this operation. However, it is not necessarily the case that higher-partitions produce faster computation; instead, partitioning data is particular to your computing cluster and the data analysis operations being performed.

## Caching {#tuning-caching}

Recall from the [introduction](Intro) that Spark was designed to be faster than its predecessors by using memory instead of disk to store data. This is formally known as a Spark **RDD** and stands for resilient distributed dataset. An RDD distributes copies of the same data across many machines, such that, if one machine fails other can complete the task -- hence the resilient name. Resiliency is important in distributed systems since, while things will usually work in one machine, when running over thousands of machines the likelihood of something failing is much higher. When a failure happens, it is preferable to be fault tolerant to avoid losing the work of all the other machines. RDDs accomplish this by tracking data lineage information to rebuild lost data automatically on failure.

In `sparklyr`, you can control when an RDD gets loaded or unloaded from memory using `tbl_cache()` and `tbl_uncache()`.

Most sparklyr operations that retrieve a Spark DataFrame, cache the results in-memory. For instance, running `spark_read_parquet()` or `sdf_copy_to()` will provide a Spark DataFrame that is already cached in-memory. As a Spark DataFrame, this object can be used in most `sparklyr` functions, including data analysis with dplyr or machine learning.

```{r tuning-caching-connect}
library(sparklyr)
sc <- spark_connect(master = "local")
```

```{r tuning-caching-copy}
iris_tbl <- sdf_copy_to(sc, iris, overwrite = TRUE)
```

You can inspect which tables are cached by navigating to the Spark UI using `spark_web(sc)`, opening the storage tab, and clicking on a given RDD:

```{r tuning-caching-rdd, echo=FALSE}
invisible(webshot::webshot(
  "http://localhost:4040/storage/rdd/?id=9",
  "images/tuning-cache-rdd-web.png",
  cliprect = "viewport"
))
```

```{r tuning-caching-rdd-shot, eval=TRUE, fig.width = 4, fig.align = 'center', echo=FALSE, fig.cap='Cached RDD in Spark Web Interface'}
render_image("images/tuning-cache-rdd-web.png", "Cached RDD in Spark Web Interface")
```

Data loaded in memory will be released when the R session terminates either explicitly or implicitly with a restart or disconnection; however, to free up resources, you can use `tbl_uncache()`:

```{r tuning-caching-uncache}
tbl_uncache(sc, "iris")
```

```{r tuning-caching-disconnect, echo=FALSE}
spark_disconnect(sc)
```

### Checkpointing

Checkpointing is a slightly different type of caching; while it also persists data it will, additionally, break the graph computation lineage. For example, if a cached partition is lost, it can be computed from the computation graph which is not possible while checkpointing since the source of computation is lost.

When performing expensive computation graphs, it can make sense to checkpoint to persist and break the computation lineage in order to help Spark reduce graph computation resources; otherwise, Spark might try to over-optimize a computation graph that is really not useful to optimize.

You can checkpoint explicitly by saving to CSV, Parquet, etc. files. Or let Spark checkpoint this for you using `sdf_checkpoint()` in `sparklyr` as follows.

Notice that checkpointing truncates the computation lineage graph, which can speed up performance if the same intermediate result is used multiple times.

### Memory {#tuning-memory}

Memory in Spark is categorized into reserved, user, execution or storage:

- **Reserved:** Reserved memory is the memory required by Spark to function and therefore, is overhead that is required and should not be configured. This value defaults to 300MB.
- **User:** User memory is the memory used to execute custom code. `sparklyr` only makes use of this memory indirectly when executing `dplyr` expressions or modeling a dataset.
- **Execution:** Execution memory is used to execute code by Spark, mostly, to process the results from the partition and perform shuffling.
- **Storage:** Storage memory is used to cache RDDs, for instance, when using `tbl_cache()` in `sparklyr`.

As part of tuning execution, you can consider tweaking the amount of memory allocated for **user**, **execution** and **storage** by creating a Spark connection with different values than the defaults provided in Spark:

```{r tuning-memory-fraction}
config <- spark_config()

# define memory available for storage and execution
config$spark.memory.fraction <- 0.75

# define memory available for storage
config$spark.memory.storageFraction <- 0.5
```

For instance, if you want to use Spark to store large amounts of data in-memory with the purpose of filtering and retrieving subsets quickly, you can expect Spark to use little execution or user memory. Therefore, to maximize storage memory, one can tune Spark as follows:

```{r tuning-memory-change}
config <- spark_config()

# define memory available for storage and execution
config$spark.memory.fraction <- 0.90

# define memory available for storage
config$spark.memory.storageFraction <- 0.90
```

However, notice that Spark will borrow execution memory from storage and viceversa if needed and if possible; therefore, in practice, there should be little need to tune the memory settings.

## Shuffling {#tuning-shuffling}

Shuffling, is the operation that redistributes data across machines; it is usually an expensive operation and therefore, one we try to minimize. One can easily identify if significant time is being spent shuffling by looking at the [event timeline](#tuning-event-timeline). It is possible to reduce shuffling by reframing data analysis questions or hinting Spark appropriately.

This would be relevant, for instance, when joining data frames that differ in size significantly, as in, one set being orders of magnitude smaller than the other one. You can consider using `sdf_broadcast()` to mark a data frame as small enough for use in broadcast joins, meaning, it pushes one of the smaller data frames to each of the worker nodes to reduce shuffling the bigger dataframe. One example for `sdf_broadcast()` follows:

```{r tuning-shuffling-boardcast}
sdf_len(sc, 10000) %>%
    sdf_broadcast() %>%
    left_join(sdf_len(sc, 100))
```

## Serialization {#tuning-serialization}

Serialization is the process of translating data and tasks into a format that can be transmitted between machines and reconstructed on the receiving end.

It is not that common to have to adjust serialization when tuning Spark; however, it is worth mentioning there are alternative serialization modules like the [Kryo Serializer](https://github.com/EsotericSoftware/kryo) that can provide performance improvements over the default [Java Serializer](http://docs.oracle.com/javase/6/docs/api/java/io/Serializable.html).

The Kryo Serializer can be enabled in `sparklyr` through:

```{r tuning-serialization-kryo}
config <- spark_config()

config$spark.serializer <- "org.apache.spark.serializer.KryoSerializer"
sc <- spark_connect(master = "local", config = config)
```

## Configuration Files

Configuring the `spark_config()` settings before connecting is the most common approach while tuning Spark. However, once the desired connection is known, you should consider switching to use a configuration file since it will remove the clutter in your connection code and also allow you to share the configuration settings across projects and coworkers.

For instance, instead of connecting to Spark through:

```{r}
config <- spark_config()
config["sparklyr.shell.driver-memory"] <- "2G"
sc <- spark_connect(master = "local", config = config)
```

You can instead define a `config.yml`  with the desired settings. This file should be located in the current working directory or in parent directories. For example, you can create the following `config.yml` file to modify the default driver memory:

```{yml}
default:
  sparklyr.shell.driver-memory: 2G
```

Then, connecting with the same configuration settings becomes much cleaner by using instead:

```{r}
sc <- spark_connect(master = "local")
```

You can also specify an alternate config file name or location by setting the `file` parameter in `spark_config()`. One additional benefit from using configuration files, is that a system administrator can change the default configuration file without modifying the original one and additional functionality provided by the `config` package, see [github.com/rstudio/config](https://github.com/rstudio/config).

## Recap

This chapter provided a broad overview of Spark internals and a detailed configuration settings to help you speed up computation and enable high computation loads. It provided the foundations to understand bottlenecks and guidance on common configuration considerations; however, fine-tuning Spark is a broad topic that would require many more chapters to cover extensively. Therefore, while troubleshooting Spark's performance and scalability, searching the web and consulting online communities it is often necessary to fine-tune your particular environment.

The next chapter, [Extensions](#extensions), introduces the ecosystem of Spark extensions available in R. Most extensions are highly-specialized, but they will prove to be extremly useful in specific cases and readers with particular needs. For instance, they can process nested data, perform graph analysis or use different modeling libraries like `rsparkling` from H20. Not only that, but the next few chapters introduce many advanced data analysis and modeling topics that are required to master large-scale computing in R.
