
# Analysis {#analysis}

## Overview {#analysis-overview}

```{r eval=FALSE, exercise=TRUE}
install.packages("ggplot2")
install.packages("corrr")
install.packages("dbplot")
install.packages("rmarkdown")
```

```{r}
library(sparklyr)
library(dplyr)

sc <- spark_connect(master = "local", version = "2.3")
```

## Import

```{r analysis-copy-to}
cars <- copy_to(sc, mtcars)
```

## Wrangle

```{r analysis-summarize-all}
summarize_all(cars, mean)
```

```{r analysis-summarize-show-query}
summarize_all(cars, mean) %>%
  show_query()
```

```{r analysis-group-summarize}
cars %>%
  mutate(transmission = ifelse(am == 0, "automatic", "manual")) %>%
  group_by(transmission) %>%
  summarise_all(mean)
```

### Built-in Functions

```{r analysis-percentile}
summarise(cars, mpg_percentile = percentile(mpg, 0.25))
```

```{r analysis-percentile-query}
summarise(cars, mpg_percentile = percentile(mpg, 0.25)) %>%
  show_query()
```

```{r analysis-percentile-rename}
summarise(cars, mpg_percentile = percentile(mpg, array(0.25, 0.5, 0.75)))
```

```{r analysis-percentile-mutate}
summarise(cars, mpg_percentile = percentile(mpg, array(0.25, 0.5, 0.75))) %>%
  mutate(mpg_percentile = explode(mpg_percentile))
```

### Correlations

```{r analysis-corrr-cars}
ml_corr(cars)
```

```{r analysis-corrr-pairwise}
library(corrr)
correlate(cars, use = "pairwise.complete.obs", method = "pearson") 
```

```{r analysis-corrr-show}
correlate(cars, use = "pairwise.complete.obs", method = "pearson") %>%
  shave() %>%
  rplot()
```

## Visualize

### Using ggplot2

```{r analysis-ggplot2-simple, eval = TRUE, out.width='500pt', out.height='400pt', fig.cap='Plotting inside R', fig.align = 'center'}
library(ggplot2)
ggplot(aes(as.factor(cyl), mpg), data = mtcars) + geom_col()
```

```{r analysis-ggplot2-simmarise}
car_group <- cars %>%
  group_by(cyl) %>%
  summarise(mpg = sum(mpg, na.rm = TRUE)) %>%
  collect() %>%
  print()
```

```{r analysis-ggplot2-groups}
ggplot(aes(as.factor(cyl), mpg), data = car_group) + 
  geom_col(fill = "#999999") + coord_flip()
```

### Using dbplot

```{r  analysis-dbplot-simple}
library(dbplot)

cars %>%
dbplot_histogram(mpg, binwidth = 3) +
labs(title = "MPG Distribution",
     subtitle = "Histogram over miles per gallon")
```

## Model

### Caching

## Communicate

```{r analysis-disconnect}
spark_disconnect(sc)
```

````markdown
---
title: "mtcars analysis"
output:
  html_document: 
    fig_width: 6 
    fig_height: 3
---
`r ''````{r, setup, include = FALSE}
library(sparklyr)
library(dplyr)

sc <- spark_connect(master = "local", version = "2.3")
cars <- copy_to(sc, mtcars)
```

## Visualize
Aggregate data in Spark, visualize in R.
`r ''````{r  fig.align='center', warning=FALSE}
library(ggplot2)
cars %>%
  group_by(cyl) %>% summarise(mpg = mean(mpg)) %>%
  ggplot(aes(cyl, mpg)) + geom_bar(stat="identity")
```

## Model
The selected model was a simple linear regression that 
uses the weight as the predictor of MPG

`r ''````{r}
cars %>%
  ml_linear_regression(wt ~ mpg) %>%
  summary()
```
`r ''````{r, include = FALSE}
spark_disconnect(sc)
```
````

## Recap
