
# Getting Started {#starting}

## Overview

## Prerequisites {#starting-prerequisites}

```{r}
system("java -version")
```

### Installing sparklyr {#starting-install-sparklyr}

```{r starting-install-sparklyr, eval=FALSE, exercise=TRUE}
install.packages("sparklyr")
```

```{r}
packageVersion("sparklyr")
```

### Installing Spark {#starting-installing-spark}

```{r starting-install-spark-header, warning=FALSE, message=FALSE}
library(sparklyr)
```

```{r starting-install-spark}
spark_install("2.3")
```

```{r starting-install-available, eval=TRUE}
spark_available_versions()
```

```{r starting-install-installed}
spark_installed_versions()
```

## Connecting {#starting-connect-to-spark}

```{r starting-connect-local}
library(sparklyr)
sc <- spark_connect(master = "local", version = "2.3")
```

## Using Spark {#starting-sparklyr-hello-world}

```{r starting-copy-cars}
cars <- copy_to(sc, mtcars)
```

```{r starting-print-cars}
cars
```

### Web Interface {#starting-spark-web-interface}

```{r starting-spark-web-code}
spark_web(sc)
```

### Analysis {#starting-analysis}

```{r}
library(DBI)
dbGetQuery(sc, "SELECT count(*) FROM mtcars")
```

```{r}
library(dplyr)
count(cars)
```

```{r}
select(cars, hp, mpg) %>%
  sample_n(100) %>%
  collect() %>%
  plot()
```

### Modeling {#starting-modeling}

```{r}
model <- ml_linear_regression(cars, mpg ~ hp)
model
```

```{r}
model %>%
  ml_predict(copy_to(sc, data.frame(hp = 250 + 10 * 1:10))) %>%
  transmute(hp = hp, mpg = prediction) %>%
  full_join(select(cars, hp, mpg)) %>%
  collect() %>%
  plot()
```

### Data {#starting-data}

```{r}
spark_write_csv(cars, "cars.csv")
```

```{r}
cars <- spark_read_csv(sc, "cars.csv")
```

### Extensions {#starting-extensions}

```{r eval=FALSE, exercise=TRUE}
install.packages("sparklyr.nested")
```

```{r}
sparklyr.nested::sdf_nest(cars, hp) %>%
  group_by(cyl) %>%
  summarise(data = collect_list(data))
```

### Distributed R {#starting-distributed-r}

```{r}
cars %>% spark_apply(~round(.x))
```

### Streaming {#starting-streaming}

```{r}
dir.create("input")
write.csv(mtcars, "input/cars_1.csv", row.names = F)
```

```{r}
stream <- stream_read_csv(sc, "input/") %>%
    select(mpg, cyl, disp) %>%
    stream_write_csv("output/")
```

```{r}
dir("output", pattern = ".csv")
```

```{r}
# Write more data into the stream source
write.csv(mtcars, "input/cars_2.csv", row.names = F)
```

```{r}
# Check the contents of the stream destination
dir("output", pattern = ".csv")
```

```{r}
stream_stop(stream)
```

### Logs {#starting-logs}

```{r starting-logs}
spark_log(sc)
```

```{r starting-logs-filter}
spark_log(sc, filter = "sparklyr")
```

## Disconnecting {#starting-disconnecting}

```{r starting-disconnect}
spark_disconnect(sc)
```

```{r starting-disconnect-all}
spark_disconnect_all()
```

## Using RStudio {#starting-using-spark-from-rstudio}

## Resources {#starting-resources}

## Recap {#starting-recap}
